---
title: "base"
output: html_document
---

```{r}
#load libraries/packages
source("code/packages.R")
library(dplyr)
library(knitr)
library(kableExtra)

# load data
#source("code/data_prep/panel_wrangling.R")
data <- readRDS("data/output/sdg_democracy_paneldata.rds")
#data <- read_csv("data/output/ES_panel_data.csv")

# save data as csv
write_csv(data, "data/output/sdg_democracy_paneldata.csv")
```

## Sumary Statistics: DI Vars: elect_dem, lib_dem, di_score
```{r}
cat("\n=== SUMMARY STATISTICS ===\n")
summary(data[, c("elect_dem", "lib_dem")])

cat("\n=== CHECK FOR MISSING VALUES ===\n")
missing_summary <- sapply(data[, c("elect_dem", "lib_dem")], function(x) sum(is.na(x)))
print(missing_summary)

cat("\n=== STRUCTURE OF THE DATA ===\n")
str(data[, c("elect_dem", "lib_dem")])
```

## Descriptive Analysis: NA Counts and DI Distributions
```{r}
# (1) Histogram: NA Counts 
hist_NA_counts <- ggplot(data, aes(x = as.numeric(prop_sdg_missing))) +
  geom_histogram(
    binwidth = 0.04,              # adjust as needed
    color = "white",
    fill = "#08519c"
  ) +
  labs(
    title = "Distribution of Missing SDG data (country-years)",
    subtitle = "NA counts, based on Sustainable Development Report (2024)",
    x = "Proportion of NA values (0-1)",
    y = "Country-year counts"
  ) +
  theme_minimal(base_size = 14)
hist_NA_counts

# (2) Histogram: EDI Score
hist_EDI_counts <- ggplot(data, aes(x = as.numeric(elect_dem))) +
  geom_histogram(
    binwidth = 0.04,              # adjust to suit scale of lib_dem
    color    = "white",
    fill     = "#238443"
  ) +
  labs(
    title = "Distribution of Democracy Levels (country-years)",
    subtitle = "Based on V-Dem's Electoral Democracy Index (EDI)",
    x     = "Electoral Democracy Levels (0-1)",
    y     = "Country-year counts"
  ) +
  theme_minimal(base_size = 14)
hist_EDI_counts

# (3) Histogram: LDI Score
hist_LDI_counts <- ggplot(data, aes(x = as.numeric(lib_dem))) +
      geom_histogram(
        binwidth = 0.04,              # adjust to suit scale of lib_dem
        color    = "white",
        fill     = "#6A51A3"
      ) +
      labs(
        title = "Distribution of Democracy Levels (country-years)",
        subtitle = "Based on V-Dem's Liberal Democracy Index (LDI)",
        x     = "Liberal Democracy Levels (0-1)",
        y     = "Country-year counts"
      ) +
      theme_minimal(base_size = 14)
hist_LDI_counts

# save all three histograms using ggsave 
ggsave(filename = "descriptive_stats/figures/hist_NA_counts.png", 
       plot = hist_NA_counts, width = 10, height = 8)
ggsave(filename = "descriptive_stats/figures/hist_EDI_counts.png", 
       plot = hist_EDI_counts, width = 10, height = 8)
ggsave(filename = "descriptive_stats/figures/hist_LDI_counts.png", 
       plot = hist_LDI_counts, width = 10, height = 8)

# combine all three histograms into a single figure using patchwork
combined_histograms <-
  hist_NA_counts  + labs(title = NULL, subtitle = NULL) +
  hist_EDI_counts + labs(title = NULL, subtitle = NULL, y = NULL) +
  hist_LDI_counts + labs(title = NULL, subtitle = NULL, y = NULL) +
  plot_annotation(
    title = "Distributions of SDG Missingness and Democracy Levels",
    caption = "Based on Sustainable Development Report (2024) and V-Dem EDI & LDI indices",
    theme = theme(
      plot.title = element_text(size = 18, face = "bold"),
      plot.subtitle = element_text(size = 14)
    )
  )

combined_histograms

ggsave(filename = "descriptive_stats/figures/combined_histograms.png", 
       plot = combined_histograms, width = 12, height = 4)
```

## Descriptive Analysis: Correlation Matrix of SDG Missingness and Democracy Levels
```{r eval=FALSE, include=FALSE}
# SDG missingness x democracy levels (elect_dem, lib_dem, di_score) 
source("descriptive_stats/corr_matrix_sdg_NAs.R")

# SDG missingness x democracy levels (elect_dem, lib_dem, di_score) [partial]
source("descriptive_stats/partial_corr_matrix_sdg_NAs.R")

# All covariates used in models
source("descriptive_stats/corr_matrix_key_vars.R")
```



## RQ1: How does SDG data missingness differ by regime type over time?
```{r}
# Ensuring regime_type_4 is labeled & ordered
data_plot <- data %>% 
  mutate(regime_type_4 = case_when(
      regime_type_4 == 0 ~ "Closed autocracy",
      regime_type_4 == 1 ~ "Electoral autocracy",
      regime_type_4 == 2 ~ "Electoral democracy",
      regime_type_4 == 3 ~ "Liberal democracy",
      TRUE ~ NA_character_),
      regime_type_4 = factor(regime_type_4,
      levels = c("Closed autocracy",
                 "Electoral autocracy",
                 "Electoral democracy",
                 "Liberal democracy")),
    
    # income_level: setting order and labels ONCE
    income_level = factor(income_level,
      levels = c("L", "LM", "UM", "H"),
      labels = c("Low income", "Lower middle", "Upper middle", "High income")),
    # Defining time periods
    period = case_when(
      year <= 2014 ~ "2000–2014",
      year <= 2018 ~ "2015–2018",
      year <= 2022 ~ "2019–2022",
      TRUE ~ "2023+"),
    
    # Venezuela-specific income variable [ONLY BECAUSE ALL NAs ARE VENEZUELA!]
    income_level_ven = case_when(
      is.na(income_level) ~ "Venezuela 2020–2023",
      TRUE ~ as.character(income_level)),
    income_level_ven = factor(
      income_level_ven,
      levels = c("Low income",
                 "Lower middle",
                 "Upper middle",
                 "High income",
                 "Venezuela 2020–2023")))

# =============================================================================
# (1) Density Plot: NA Counts by RoW Groups (Regime Types)
dens_plt_regimetype <- ggplot(data_plot, 
             aes(x = prop_sdg_missing, 
                 fill = regime_type_4, 
                 color = regime_type_4)) +
        geom_density(alpha = 0.30, adjust = 1) +
        labs(title = "Distribution of Missing SDG data, by Regime Type",
          caption = "Based on Sustainable Development Report (2024) and V-Dem's RoW classifications",
          x = "Proportion of SDG indicator values that are NA (per country-year)",
          y = "Density",
          fill = "Regime type",
          color = "Regime type") +
        theme_minimal(base_size = 20) +
  theme(legend.position = "top", legend.box = "horizontal")
dens_plt_regimetype

# saving density plot 1 
ggsave(filename = "descriptive_stats/figures/dens_plt_regimetype.png", 
       plot = dens_plt_regimetype, width = 15, height = 8)

# sanity check 
table(data_plot$income_level, useNA = "ifany")
table(data_plot$income_level_ven, useNA = "ifany")

# =============================================================================
# (2) Lineplot: NA Counts by RoW Groups over Time (Regime Types)
year_regime_means <- data_plot %>% 
  group_by(year, regime_type_4) %>% 
  summarise(mean_missing = mean(prop_sdg_missing),
    .groups = "drop")

line_plt_regxtime <- ggplot(year_regime_means,
             aes(x = year,
                 y = mean_missing,
                 color = regime_type_4)) +
        geom_line(size = 1) +
        geom_point(size = 1.5) +
        labs(
          title = "SDG missingness by regime type over time",
          subtitle = "Measured in mean proportion of NA values by regime type, per year",
          x = NULL,
          y = "Mean proportion of NA values",
          color = "Regime type",
          caption = "Based on Sustainable Development Report (2024) and V-Dem's RoW classifications"
        ) +
        theme_minimal(base_size = 20) +
  theme(legend.position = "top", legend.box = "horizontal")
line_plt_regxtime

# =============================================================================
# (3) Boxplot: NA Counts by RoW Groups
box_plt_regxtime <- ggplot(data_plot,
             aes(x = factor(year),
                 y = prop_sdg_missing)) +
        geom_boxplot(outlier.alpha = 0.2) +
        facet_wrap(~ regime_type_4, nrow = 2) +
        labs(
          title = "Distribution of SDG missingness by year and regime type",
          x = NULL,
          y = "Proportion of SDG indicator values that are NA") +
        theme_minimal(base_size = 11) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
box_plt_regxtime

# saving line and box plots 2 & 3
ggsave(filename = "descriptive_stats/figures/line_plt_regxtime.png", 
       plot = line_plt_regxtime, width = 15, height = 8)
ggsave(filename = "descriptive_stats/figures/box_plt_regxtime.png", 
       plot = box_plt_regxtime, width = 15, height = 8)

# combine dens_plt_regimetype and line_plt_regxtime into a single figure using patchwork

library(patchwork)

combined_regime_plots <- (dens_plt_regimetype /
      labs(title = NULL, subtitle = NULL, caption = NULL, x = NULL) +
      theme(legend.position = "top", legend.justification = c(0.1, 1)) + 
    line_plt_regxtime +
      labs(title = NULL, subtitle = NULL) +
      theme(legend.position = "none")) + 
  plot_annotation(
    title = "SDG Missingness by Regime Type",
    subtitle = "Comparing distributions and trends over time",
    theme = theme(
      plot.title    = element_text(size = 40, face = "bold"),
      plot.subtitle = element_text(size = 33)))

combined_regime_plots

ggsave(
  filename = "descriptive_stats/figures/combined_regime_plots.png",
  plot     = combined_regime_plots,
  width    = 15,
  height   = 12
)
```

## Regime Change Status: Comparing Backsliding Years to All Other Country-Year Types

## RQ2: How does SDG data missingness differ between country–years experiencing democratic backsliding and all other country–years?
- Group 1: all country–years where aut_ep = 1
- Group 2: all country–years where aut_ep = 0 (including always-stable democracies, always-autocracies, and pre/post years of backsliders).
```{r}
# BACKSLIDING VS ALL ELSE 
means_by_backsliding <- data %>% 
  group_by(aut_ep) %>% 
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    se_missing = sd(prop_sdg_missing, na.rm = TRUE) / sqrt(n()),
    n = n())

# BAR PLOT: BACKSLIDING VS ALL ELSE 
bar_plt_back_nonback <-ggplot(means_by_backsliding, 
                              aes(x = factor(aut_ep, levels = c(0, 1),
                            labels = c("All else", "Backsliding")),
                 y = mean_missing,
                 fill = factor(aut_ep))) +
        geom_col(width = 0.70) +
        geom_errorbar(aes(ymin = mean_missing - se_missing,
                          ymax = mean_missing + se_missing),
                      width = 0.15) + 
  labs(title = "Average SDG data missingness",
       subtitle = "Backsliding vs non-backsliding years (all country-years)",
       x = NULL, 
       y = "Missingness rate", 
       caption = "Based on V-Dem's Episodes of Regime Transition (ERT) Data") +
  theme_minimal() +
  theme(legend.position = "none")
bar_plt_back_nonback

# save bar plot
ggsave(filename = "descriptive_stats/figures/bar_plt_back_nonback.png", 
       plot = bar_plt_back_nonback, width = 12, height = 8)
```

## BACKSLIDING VS DEMOCRATIZING VS STABLE
```{r}
# BACKSLIDING VS DEMOCRATIZING VS STABLE
regime_data <- data %>%
  mutate(regime_status = case_when(
      aut_ep == 1 ~ "Backsliding",
      dem_ep == 1 ~ "Democratizing",
      TRUE ~ "Stable"))

means_by_status <- regime_data %>%
  group_by(regime_status) %>%
  summarise(mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    se_missing = sd(prop_sdg_missing, na.rm = TRUE) / sqrt(n()),
    n = n())

# BAR PLOT: BACKSLIDING VS DEMOCRATIZING VS STABLE
bar_plt_back_dem_stab <- ggplot(means_by_status, aes(x = regime_status,
                            y = mean_missing,
                            fill = regime_status)) +
  geom_col(width = 0.65) +
  geom_errorbar(aes(ymin = mean_missing - 1.96 * se_missing,
                    ymax = mean_missing + 1.96 * se_missing),
                width = 0.15) +
  labs(title = "Average SDG data missingness by transition type",
       subtitle = "Backsliding vs democratizing vs stable periods (all country-years)",
       x = NULL,
       y = "Missingness rate",
       caption = "Based on V-Dem's Episodes of Regime Transition (ERT) Data") +
  theme_minimal() +
  theme(legend.position = "none")
bar_plt_back_dem_stab

# save bar plot
ggsave(filename = "descriptive_stats/figures/bar_plt_back_dem_stab.png", 
       plot = bar_plt_back_dem_stab, width = 12, height = 8)


# COMBINED BAR PLOTS: BACKSLIDING VS DEMOCRATIZING VS STABLE + BACKSLIDING VS ALL ELSE
combined_bars <-
  bar_plt_back_nonback + labs(title = NULL, caption = NULL) +
  bar_plt_back_dem_stab + labs(title = NULL) +
  plot_layout(
    ncol = 1,          # 2 rows, 1 column (stacked vertically)
    heights = c(1, 1), # equal height; adjust if you want
    guides = "collect" # pulls the legend into a shared guide area
  ) +
  plot_annotation(
    title = "Average SDG data missingness by regime status",
    subtitle = "Comparing backsliding/non-backsliding years and transition types",
    theme = theme(
      plot.title  = element_text(size = 18, face = "bold"),
      plot.subtitle = element_text(size = 14))) &
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.x = element_text(size = 14))

combined_bars

ggsave(filename = "descriptive_stats/figures/combined_bars.png", 
       plot = combined_bars, width = 12, height = 10)

# SUMMARY TABLE 
reg_status_sum_table <- regime_data %>%
  group_by(regime_status) %>%
  summarise(
    n_obs = n(),
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    sd_missing = sd(prop_sdg_missing,   na.rm = TRUE),
    se_missing = sd_missing / sqrt(n_obs),
    ci_lower = mean_missing - 1.96 * se_missing,
    ci_upper = mean_missing + 1.96 * se_missing,
    .groups = "drop")

# formatting table
reg_status_sum_table <- reg_status_sum_table %>%
  kable(digits = 3, caption = "Summary statistics of SDG missingness by regime status") %>%
  kable_classic_2(full_width = FALSE, position = "center")

# save table
save_kable(reg_status_sum_table, "descriptive_stats/tables/reg_status_sum_table.png")
```

## BACKSLIDING VS ALL ELSE: LAGGED AND LEAD MISSINGNESS
```{r}
# BACKSLIDING VS ALL ELSE: LAGGED AND LEAD MISSINGNESS
regime_data <- data %>%
  mutate(regime_status = case_when(
      aut_ep == 1 ~ "Backsliding",
      dem_ep == 1 ~ "Democratizing",
      TRUE ~ "Stable"))

regime_data <- regime_data %>%
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    missing_lag1 = dplyr::lag(prop_sdg_missing, 1),  # year t-1
    missing_lead1 = dplyr::lead(prop_sdg_missing, 1)  # year t+1
  ) %>%
  ungroup()

# Summary: keep only backsliding years at t-1, t, t+1
back_lag_lead_means <- regime_data %>%
  filter(aut_ep == 1) %>%
  summarise(
    n_obs = n(),
    group = "Backsliding years",
    mean_tminus1 = mean(missing_lag1,  na.rm = TRUE),
    mean_t = mean(prop_sdg_missing,  na.rm = TRUE),
    mean_tplus1 = mean(missing_lead1, na.rm = TRUE)
  )
back_lag_lead_means

# Summary: Non-backsliding years at t-1, t, t+1
non_back_lag_lead_means <- regime_data %>%
  filter(aut_ep == 0) %>%
  summarise(
    n_obs = n(),
    group = "Non-backsliding years",
    mean_tminus1 = mean(missing_lag1,  na.rm = TRUE),
    mean_t = mean(prop_sdg_missing,  na.rm = TRUE),
    mean_tplus1  = mean(missing_lead1, na.rm = TRUE)
  )
non_back_lag_lead_means

# combines back_lag_lead_means & non_back_lag_lead_means
lag_lead_compare <- bind_rows(back_lag_lead_means, non_back_lag_lead_means)
lag_lead_compare
```
**INTERPRETATION:** Across 348 backsliding country–years, average SDG missingness is about 0.461 in the year before, 0.483 in the backsliding year, and 0.487 in the year after.


## RQ3: Among countries that experience democratic backsliding, how does SDG data missingness in backsliding years compare to their own non‑backsliding years? [BACKSLIDING COUNTRIES ONLY]

```{r}
# BACKSLIDING YEARS VS NON-BACKSLIDING YEARS (BACKSLIDING COUNTRIES ONLY)
# Flagging backsliding countries
backsliders <- regime_data %>%
  group_by(country_code) %>%
  summarise(ever_backslid = any(aut_ep == 1, na.rm = TRUE),
            .groups = "drop") %>%
  filter(ever_backslid) %>%
  pull(country_code)

# Filtering out all other countries 
regime_back_countries <- regime_data %>%
  filter(country_code %in% backsliders) %>%
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(backslide_status = ifelse(aut_ep == 1,
                              "Backsliding years",
                              "Non-backsliding years"),
    missing_lag1  = dplyr::lag(prop_sdg_missing, 1),  # t-1
    missing_lead1 = dplyr::lead(prop_sdg_missing, 1)  # t+1
  ) %>%
  ungroup()
# =============================================================================
combined_summary <- regime_back_countries %>%
  mutate(group = if_else(aut_ep == 1,
                         "Backsliding years",
                         "Non-backsliding years")) %>%
  group_by(group) %>%
  summarise(
    n_obs = n(),
    
    # t-level
    mean_t   = mean(prop_sdg_missing, na.rm = TRUE),
    median_t = median(prop_sdg_missing, na.rm = TRUE),
    sd_t     = sd(prop_sdg_missing,   na.rm = TRUE),
    se_t     = sd_t / sqrt(n_obs),
    ci_t_lo  = mean_t - 1.96 * se_t,
    ci_t_hi  = mean_t + 1.96 * se_t,

    # t-1
    mean_tminus1   = mean(missing_lag1, na.rm = TRUE),
    median_tminus1 = median(missing_lag1, na.rm = TRUE),
    sd_tminus1     = sd(missing_lag1,   na.rm = TRUE),
    se_tminus1     = sd_tminus1 / sqrt(n_obs),

    # t+1
    mean_tplus1   = mean(missing_lead1, na.rm = TRUE),
    median_tplus1 = median(missing_lead1, na.rm = TRUE),
    sd_tplus1     = sd(missing_lead1,   na.rm = TRUE),
    se_tplus1     = sd_tplus1 / sqrt(n_obs),

    .groups = "drop"
  )

# COMPACTING TABLE FOR PRESENTATION
combined_long <- combined_summary %>%
  select(group, n_obs,
         mean_tminus1,  median_tminus1,  se_tminus1,
         mean_t,        median_t,        se_t,
         mean_tplus1,   median_tplus1,   se_tplus1) %>%
  pivot_longer(
    cols = starts_with("mean_") | starts_with("median_"),
    names_to  = c("stat", "time"),
    names_pattern = "(mean|median)_t?(minus1|plus1)?",
    values_to = "value"
  ) %>%
  mutate(
    time = case_when(
      time == "minus1" ~ "t-1",
      time == ""       ~ "t",
      time == "plus1"  ~ "t+1"
    )
  ) %>%
  pivot_wider(
    names_from  = stat,
    values_from = value
  ) %>%
  mutate(
    se = case_when(
      time == "t-1" ~ se_tminus1,
      time == "t"   ~ se_t,
      time == "t+1" ~ se_tplus1
    )
  ) %>%
  select(group, time, n_obs, mean, median, se)

# FORMATTING TABLE 
combined_long_tab <- combined_long %>%
  kable(
    digits = 3,
    col.names = c("Group", "Time", "N",
                  "Mean missingness", "Median missingness", "SE"),
    caption = "Average and median SDG missingness in t−1, t, and t+1 for backsliding vs non-backsliding years (backsliding countries only)"
  ) %>%
  kable_classic_2(full_width = FALSE, position = "center")

combined_long_tab

# SAVE 
save_kable(combined_long_tab, "descriptive_stats/tables/rq2_summary_tab.png")

# =============================================================================
# BAR PLOT: BACKSLIDING YEARS VS NON-BACKSLIDING YEARS (BACKSLIDING COUNTRIES ONLY)
bar_plt_back_nonback_back_only <- ggplot(combined_summary, 
                                         aes(x = group, 
                                             y = mean_t,
                                             fill = group)) +
  geom_col(width = 0.70) +
  geom_errorbar(aes(ymin = ci_t_lo, ymax = ci_t_hi), width = 0.15) +
  labs(
    title = "Average SDG data missingness",
    subtitle = "Backsliding vs non-backsliding years (backsliding countries only)",
    x = NULL,
    y = "Average SDG missing rate",
    caption = "Based on V-Dem's ERT Autocratization Episodes"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
bar_plt_back_nonback_back_only

#save 
ggsave(filename = "descriptive_stats/figures/bar_plt_back_nonback_back_only.png", 
       plot = bar_plt_back_nonback_back_only, width = 12, height = 8)
```

## RQ4: Across all country-years, what is the distribution of missingness by income group?
Income Groups & Liberal Democracy Analysis (ld_data): NA Counts x Year x Income Level
```{r}
# (4) Density Plot: NA Counts by Income Groups
dens_plt_income <- ggplot(data_plot, 
             aes(x = prop_sdg_missing, 
                 fill = income_level_ven, 
                 color = income_level_ven)) +
        geom_density(alpha = 0.30, adjust = 1) +
        labs(title = "Distribution of Missing SDG data, by Income Level",
          caption = "Based on Sustainable Development Report (2024) and World Bank Gross National Income (GNI) groupings",
          x = "Proportion of SDG indicator values that are NA (per country-year)",
          y = "Density",
          fill = "Income level",
          color = "Income level") +
        theme_minimal(base_size = 12)
dens_plt_income

# saving plot X
ggsave(filename = "descriptive_stats/figures/dens_plt_income.png", 
       plot = dens_plt_income, width = 12, height = 8)

# =============================================================================
# Focusing on RoW Liberal Democracy
ld_data <- data_plot %>%
  filter(regime_type_4 == "Liberal democracy")

# (5) Boxplot of full distribution by year (within liberal democracies)
box_plt_libdemxtime <- ggplot(ld_data, aes(x = factor(year), y = prop_sdg_missing)) +
        geom_boxplot(outlier.alpha = 0.2) +
        labs(
          title = "Distribution of SDG missingness by year",
          subtitle = "(Liberal democracies only)",
          x = NULL,
          y = "Missigness rate") +
        theme_minimal(base_size = 12) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
box_plt_libdemxtime

# saving plot 5
ggsave(filename = "descriptive_stats/figures/box_plt_libdemxtime.png", 
       plot = box_plt_libdemxtime, width = 12, height = 8)

# =============================================================================
# (6) Density by income (liberal democracies only)
dens_plt_income_libdem <- ggplot(ld_data,
             aes(x = prop_sdg_missing,
                 fill  = income_level,
                 color = income_level)) +
        geom_density(alpha = 0.30, adjust = 1) +
        labs(title = "Distribution of Missing SDG data, by Income Group",
             subtitle = "(Liberal democracies only)",
          x = "Missigness rate",
          y = "Density",
          fill = "Income level",
          color = "Income level", 
          caption = "Based on Sustainable Development Report (2024) and RoW Regime type") +
        theme_minimal(base_size = 12)
dens_plt_income_libdem

# saving plot 6
ggsave(filename = "descriptive_stats/figures/dens_plt_income_libdem.png", 
       plot = dens_plt_income_libdem, width = 12, height = 8)

# Quick sanity check
table(ld_data$period, useNA = "ifany")

# =============================================================================
# (7) Boxplot of full distribution by year, faceted by income level (within liberal democracies)
box_plt_libdemxtimexincome <- ggplot(ld_data, aes(x = factor(year), y = prop_sdg_missing)) +
  geom_boxplot(outlier.alpha = 0.2) +
  facet_wrap(~ income_level, nrow = 1) +
  labs(title = "Distribution of SDG missingness by income level over time",
  subtitle = "(Liberal democracies only)",
    x = NULL,
    y = "Missingness rate", 
  caption = "Subset based on V-Dem's RoW classifications") +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
box_plt_libdemxtimexincome

ggsave(filename = "descriptive_stats/figures/box_plt_libdemxtimexincome.png", 
       plot = box_plt_libdemxtimexincome, width = 12, height = 8)

# =============================================================================
# (8) Lineplot of mean missingness by year, faceted by income level (within liberal democracies)
year_income_means <- ld_data %>% 
  group_by(year, income_level) %>% 
  summarise(mean_missing = mean(prop_sdg_missing), .groups = "drop")

line_plt_income_libdem <- ggplot(year_income_means, 
       aes(x = year, 
           y = mean_missing,
           color = income_level)) + 
  geom_line(size = 1) + 
  geom_point(size = 1.5) +
  labs(title = "Average SDG missingness over time by income level", 
  subtitle = "(Liberal democracies only)",
    x = NULL,
    y = "Missigness rate",
    color = "Income level") + 
  theme_minimal(base_size = 13)
line_plt_income_libdem

ggsave(filename = "descriptive_stats/figures/line_plt_income_libdem.png", 
       plot = line_plt_income_libdem, width = 12, height = 8)
```


## RQ5: Which SDG goals show higher average data missingness in autocratizing years than in non‑autocratizing years, and how large are these goal‑specific gaps?
```{r}
regime_long <- regime_data %>%
  pivot_longer(
    cols = starts_with("prop_miss_SDG"),
    names_to = "sdg_goal",
    values_to = "prop_sdg_missing_goal"
  ) %>%
  mutate(sdg_goal = gsub("prop_miss_", "", sdg_goal)  # "SDG1", "SDG2", ...
  )

goal_means <- regime_long %>%
  group_by(sdg_goal, aut_ep) %>%        # aut_ep: 1 = autocratizing year, 0 = other
  summarise(
    n_obs        = n(),
    mean_missing = mean(prop_sdg_missing_goal, na.rm = TRUE),
    sd_missing   = sd(prop_sdg_missing_goal,   na.rm = TRUE),
    se_missing   = sd_missing / sqrt(n_obs),
    .groups      = "drop_last"
  )

goal_diff_table <- goal_means %>%
  mutate(aut_label = ifelse(aut_ep == 1, "aut", "nonaut")) %>%
  select(-aut_ep) %>%
  pivot_wider(
    names_from  = aut_label,
    values_from = c(n_obs, mean_missing, sd_missing, se_missing),
    names_sep   = "_"
  ) %>%
  mutate(
    diff_mean = mean_missing_aut - mean_missing_nonaut,
    se_diff   = sqrt(se_missing_aut^2 + se_missing_nonaut^2),
    ci_lower  = diff_mean - 1.96 * se_diff,
    ci_upper  = diff_mean + 1.96 * se_diff
  ) %>%
  arrange(desc(diff_mean))   # goals with largest gaps at top

goal_diff_table

# =============================================================================
# TABLE PREP 

# SDG labels lookup
sdg_labels <- tibble::tibble(
  sdg_goal = paste0("SDG", 1:17),
  sdg_label = c(
    "No poverty",
    "Zero hunger",
    "Good health & well-being",
    "Quality education",
    "Gender equality",
    "Clean water & sanitation",
    "Affordable & clean energy",
    "Decent work & economic growth",
    "Industry, innovation & infrastructure",
    "Reduced inequalities",
    "Sustainable cities & communities",
    "Responsible consumption & production",
    "Climate action",
    "Life below water",
    "Life on land",
    "Peace, justice & strong institutions",
    "Partnerships for the goals"
  )
)

goal_diff_clean <- goal_diff_table %>%
  # join labels
  left_join(sdg_labels, by = "sdg_goal") %>%
  # extract numeric SDG number for proper ordering
  mutate(sdg_num = as.integer(gsub("SDG", "", sdg_goal))) %>%
  select(
    sdg_num,
    sdg_goal,
    sdg_label,
    n_obs_nonaut,
    n_obs_aut,
    mean_missing_nonaut,
    mean_missing_aut,
    diff_mean,
    ci_lower,
    ci_upper
  ) %>%
  arrange(desc(diff_mean)) %>%  # or arrange(sdg_num) for 1–17 order
  mutate(
    across(
      c(mean_missing_nonaut, mean_missing_aut, diff_mean, ci_lower, ci_upper),
      ~ round(.x, 3)
    )
  )

goal_diff_kable <- goal_diff_clean %>%
  select(
    sdg_goal,
    sdg_label,
    n_obs_nonaut,
    n_obs_aut,
    mean_missing_nonaut,
    mean_missing_aut,
    diff_mean,
    ci_lower,
    ci_upper
  ) %>%
  kable(
    format = "html",
    caption = "Table X. Average SDG missingness in autocratizing vs non-autocratizing years by goal",
    col.names = c(
      "SDG",
      "Goal label",
      "N (non-aut.)",
      "N (aut.)",
      "Mean non-aut.",
      "Mean aut.",
      "Diff (aut − non-aut)",
      "95% CI lower",
      "95% CI upper"
    ),
    align = "llrrrrrrr"
  ) %>%
  kable_classic_2(full_width = FALSE)
goal_diff_kable

save_kable(goal_diff_kable, "descriptive_stats/tables/goal_diff_table.png")

# =============================================================================
# ERROR-BAR PLOTTING!
# Settup for plotting
goal_diff_plot_data <- goal_diff_clean %>%
  mutate(
    sdg_full = paste0(sdg_goal, " ", sdg_label),
    # reverse so largest diff at top .. [NOT WORKING]
    sdg_full = factor(sdg_full, levels = rev(sdg_full))) 

erbar_plt_goals <- ggplot(goal_diff_plot_data,
                          aes(x = diff_mean, y = sdg_full)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
  geom_pointrange(aes(xmin = ci_lower, xmax = ci_upper)) +
  labs(
    x = "Difference in mean missingness (aut − non-aut)",
    y = NULL,
    title = "Goal-specific gaps in SDG missingness,\nautocratizing vs non-autocratizing years"
  ) +
  theme_minimal()

erbar_plt_goals

# saving 
ggsave(filename = "descriptive_stats/figures/erbar_plt_goals.png", 
       plot = erbar_plt_goals, width = 12, height = 8)
```

## RQX: OTHER ANALYSES
```{r}
# Within-country variation in missingness
sdg_within <- data %>%
  group_by(country_code) %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    sd_missing = sd(prop_sdg_missing, na.rm = TRUE),
    mean_log_gdppc = mean(log_gdppc, na.rm = TRUE),
    sd_log_gdppc = sd(log_gdppc, na.rm = TRUE),
    mean_lib_dem = mean(lib_dem, na.rm = TRUE),
    sd_lib_dem = sd(lib_dem, na.rm = TRUE),
    # income level that appears most frequently for that country (mode)
    income_level_mode = names(sort(table(income_level), decreasing = TRUE))[1],
    # mode of regime type for that country
    regime_type_mode = names(sort(table(regime_type_4), decreasing = TRUE))[1],
    n_years = n()
  ) %>%
  arrange(desc(mean_missing))

# scatterplot of mean missingness vs. gdp per capita, colored by regime type
ggplot(sdg_within, aes(x = mean_log_gdppc, y = mean_missing)) +
  geom_point(aes(color = regime_type_mode), size = 1) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Within-Country Variation in SDG Data Missingness",
    x = "Average log GDP per capita (2015–2023)", 
    y = "Mean Missingness Rate (2015–2023)",
    caption = "Source: Sustainable Development Report (2025)"
  )
```

## RQX: The number of countries experiencing backsliding episodes over time by income group
```{r}
# Number of backsliding episodes by year and income group
backsliding_trends <- data %>%
  filter(aut_ep == 1) %>%
  group_by(year, income_level) %>%
  summarise(n_backsliding = n(), .groups = "drop")
# Line plot of backsliding episodes over time by income group
ggplot(backsliding_trends, aes(x = year, y = n_backsliding, color = income_level)) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  labs(
    title = "Trends in Democratic Backsliding Episodes by Income Group",
    x = NULL,
    y = "Number of backsliding episodes",
    color = "Income group",
    caption = "Based on V-Dem's Episodes of Regime Transition (ERT) Data"
  ) +
  theme_minimal(base_size = 13)

```

## how many countries have experienced pre-backsliding years (aut_ep_pre_yr == 1) 
```{r}
pre_backsliding_counts <- data %>%
  filter(aut_ep_pre_yr == 1) %>%
  group_by(year, income_level) %>%
  summarise(n_pre_backsliding = n(), .groups = "drop")
pre_backsliding_counts

# bargraph of pre-backsliding years by income group
ggplot(pre_backsliding_counts, aes(x = year, y = n_pre_backsliding, fill = income_level)) +
  geom_col(position = "dodge") +
  labs(
    title = "Number of Pre-Backsliding Years by Income Group",
    x = NULL,
    y = "Number of pre-backsliding years",
    fill = "Income group",
    caption = "Based on V-Dem's Episodes of Regime Transition (ERT) Data"
  ) +
  theme_minimal(base_size = 13)
```


## Trend Analysis 
```{r}
# load function 
source("code/SDGCountryTrends_Function.R")

create_sdg_trends_plot() # (country_name, sdg_numbers, data, sdg_colors)
```




