---
title: 'DSPC 7514 Deliverable 2: \break Project Proposal (SDG Analysis)'
author: "Priskila Nandita (ptn2113) , Naz Hussein (nh2845), Sevastian Sanchez (ss7257)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    highlight: tango
    theme: default
    fig_caption: yes
    df_print: tibble
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load necessary libraries
```{r}
library(tidyverse)
library(readxl)
library(devtools)
library(vdemdata) # call vdem package 
library(ERT) # call ERT package
library(WDI) # call WDI package for GINI coefficient
library(janitor)
library(knitr)
```

## Loading SDG Data
*Using most recent SDG data from the Sustainable Development Report 2025*
```{r}
# Filter-out pseudo countries: 
ids_to_drop <- c(
  "_Africa", "_BRICS", "_BRICSplus", "_E_Euro_Asia", "_E_S_Asia", "_HIC",
  "_LAC", "_LIC", "_LIC_LMIC", "_LMIC", "_MENA", "_OECD", "_Oceania",
  "_SIDS", "_UMIC", "_World", "_LAC"
)

# sdg excel from directory (Raw indicators)
sdg_raw <- read_excel("data/input/SDR2025-data.xlsx", sheet = "All Raw Data") %>%
  filter(year >= 2000) %>% 
  filter(!id %in% ids_to_drop)
```

## SDG Data Prep (reshaping from wide to long format)
```{r}
sdg_long <- sdg_raw %>%
  pivot_longer(
    cols = starts_with("sdg"),
    names_to  = "code",
    values_to = "value"
  ) %>%
  mutate(value = suppressWarnings(as.numeric(value)))  # in case of text like ".."
```


# 1) Creating country–year missingness variable
**Time period: 2000–2025 (SDG era: 2015–2025)**
```{r}
# CREATING MISSINGNESS VARIABLE 
sdg_missing_cy <- sdg_long %>%
  group_by(Country, year) %>%
  summarise(
    total_indicators = n(),
    missing_indicators = sum(is.na(value)),
    missing_rate = missing_indicators / total_indicators, # VARIABLE OF INTEREST
    .groups = "drop"
  )

# SUMMARIZING MISSINGESS VARIABLE
sdg_summary <- sdg_missing_cy %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    min_missing = min(missing_rate, na.rm = TRUE),
    max_missing = max(missing_rate, na.rm = TRUE),
    n_countries = n_distinct(Country),
    n_obs = n()
  )
sdg_summary

kable(sdg_summary, caption = "Summary Statistics: SDG Data Missingness (2000–2025), country-years")
```

**Interpretation** 
On average, `r round(sdg_summary$mean_missing * 100, 1)`% of SDG indicators 
are missing for a given country in a given year — meaning countries, on average,
report only about `r round((1 - sdg_summary$mean_missing) * 100, 1)` of 
SDG indicators.

The best-reported country-year had only `r round(sdg_summary$min_missing * 100, 1)`
% missing, while the worst had `r round(sdg_summary$max_missing * 100, 1)`%.

There’s moderate variation across countries and years — some countries are 
much more complete than others. The data has a standard deviation of 
`r round(sdg_summary$sd_missing * 100, 1)` percentage points.

The data includes `r round(sdg_summary$n_countries)` countries or territories, 
consistent with global SDG coverage.

# 2) Missigness (Variation) Across Time 
```{r}
# summarizing missingness by year
sdg_missing_y <- sdg_missing_cy %>%
  group_by(year) %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    median_missing = median(missing_rate, na.rm = TRUE), 
    n_countries = n_distinct(Country))

kable(sdg_missing_y, caption = "Summary Statistics: SDG Data Missingness (2000–2025), yearly stats")

# Missingness over time line graph
ggplot(sdg_missing_y, aes(x = year, y = mean_missing)) +
  geom_line(color = "#2C3E50", size = 0.8) +
  geom_point(color = "#E74C3C", size = 1) +
  geom_ribbon(aes(ymin = mean_missing - sd_missing, ymax = mean_missing + sd_missing),
              fill = "skyblue", alpha = 0.2) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Average SDG Data Missingness Over Time (2000–2024)",
    x = "Year",
    y = "Mean Missing Rate",
    caption = "Source: Sustainable Development Report (2025)")

```

**Interpretation**
The Average SDG Data Missingness Over Time graph shows the trend in the global SDG data missingness rate from 2000 to 2025. Between 2000 and 2019, the global average missingness steadily declined from roughly 60% to below 45% in 2019, reflecting improvements in data coverage and reporting consistency under the early SDG framework.

However, after 2019, the trend reverses sharply. By 2025, the global mean missingness reaches `r round(max(sdg_missing_y$mean_missing, na.rm = TRUE) * 100, 1)`%, 
indicating that only about `r round((1 - max(sdg_missing_y$mean_missing, na.rm = TRUE)) * 100, 1)` % of SDG indicators are being reported on average. 

The relative increase in missingness between 2019 and 2021 is particularly striking, with the mean missingness rate jumping from around `r round(sdg_missing_y$mean_missing[sdg_missing_y$year == 2019] * 100, 1)`% to `r round(sdg_missing_y$mean_missing[sdg_missing_y$year == 2021] * 100, 1)`%, suggesting a significant disruption in SDG data reporting, which coincides with post-pandemic disruptions, capacity shortfalls, and possible political restrictions on data transparency.

Across the entire 2015–2024 period, countries report an average of 
`r round((1 - sdg_summary$mean_missing) * 100, 1)`% of SDG indicators, with a standard deviation of `r round(sdg_summary$sd_missing * 100, 1)` percentage points based on `r scales::comma(sdg_summary$n_obs)` country-year observations covering `r sdg_summary$n_countries` countries. The lowest level is around `r round(min(sdg_missing_y$mean_missing, na.rm = TRUE) * 100, 1)`% in `r sdg_missing_y$year[which.min(sdg_missing_y$mean_missing)]` and the highest reaching `r round(max(sdg_missing_y$mean_missing, na.rm = TRUE) * 100, 1)`% in `r sdg_missing_y$year[which.max(sdg_missing_y$mean_missing)]`.  

Overall, the figure suggests that initial progress toward comprehensive SDG 
reporting has been followed by a dramatic reversal, which may not necessarily reflect a decline in reporting but a lag in data collection and processing. This overall trend cannot alone provide evidence to suggest a systemic decline in reporting capacity because it can take time for data to be collected, processed, and published, especially in the wake of global disruptions.


# 3) Cross-Sectional Distribution (2020 snapshot)
```{r}
sdg_missing_cy %>%
  filter(year == 2020) %>%
  ggplot(aes(x = missing_rate)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  theme_bw() +
  labs(title = "Distribution of SDG Data Missingness (2020)",
       x = "Missing Rate", y = "Number of Countries")
```
       
**Interpretation**
The Distribution of SDG Data Missingness graph shows the distribution of SDG 
data missingness across countries in 2020. The average missingness rate that year was `r round(mean(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)` %, with a standard deviation of `r round(sd(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)` percentage points.  

Most countries reported between `r round(min(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)`% and `r round(max(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)`% of their SDG indicators, indicating moderate variation in reporting completeness 
across countries.  

The histogram shows a concentration around the mid-range (roughly 40–50%), 
suggesting that while many countries achieved partial reporting, 
full SDG data coverage remained uncommon in 2020.


# 4) Within-Country Variation
```{r}
sdg_within <- sdg_missing_cy %>%
  group_by(Country) %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    n_years = n()
  ) %>%
  arrange(desc(mean_missing))

# scatterplot of mean missingness vs. within-country variation
ggplot(sdg_within, aes(x = sd_missing, y = mean_missing)) +
  geom_point(color = "#2980B9", size = 2) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Within-Country Variation in SDG Data Missingness",
    x = "Standard Deviation of Missing Rate", 
    y = "Mean Missing Rate (2000–2025)",
    caption = "Source: Sustainable Development Report (2025)"
  )

```


# Ensuring pseudo countries are filtered across datasets
```{r eval=FALSE, include=FALSE}
# Filter-out pseudo countries: 
ids_to_drop <- c(
  "_Africa", "_BRICS", "_BRICSplus", "_E_Euro_Asia", "_E_S_Asia", "_HIC",
  "_LAC", "_LIC", "_LIC_LMIC", "_LMIC", "_MENA", "_OECD", "_Oceania",
  "_SIDS", "_UMIC", "_World", "_LAC")

# sdg excel from directory (composite scores)
sdg <- read_excel("data/input/SDR2025-data.xlsx", sheet = "Backdated SDG Index") %>%
  filter(year >= 2000) %>% 
  filter(!id %in% ids_to_drop)

# panel data (for income levels)
ES_panel_data <- read_csv("data/output/ES_panel_data.csv") %>%
  filter(year >= 2000) %>% 
  filter(!country_code %in% ids_to_drop)
```
