---
title: 'DSPC 7514 Deliverable 2: \break Project Proposal (SDG Analysis)'
author: "Priskila Nandita (ptn2113) , Naz Hussein (nh2845), Sevastian Sanchez (ss7257)"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    highlight: tango
    theme: default
    fig_caption: yes
    df_print: tibble
urlcolor: blue
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Load necessary libraries
```{r}
library(tidyverse)
library(readxl)
library(devtools)
library(vdemdata) # call vdem package 
library(ERT) # call ERT package
library(WDI) # call WDI package for GINI coefficient
library(janitor)
library(knitr)
```

# set working directory 

#sdg excel from directory (Raw indicators)
```{r}
setwd("~/Desktop/SIPA/Fall 25' - SIPA/Policy Data Analysis Using R/DataLoss_Proj")

sdg_raw <- read_excel("data/SDR2025-data.xlsx", sheet = "All Raw Data") %>%
  filter(year >= 2000)
view(sdg_raw)
```

#sdg excel from directory (composite scores)
```{r}
sdg <- read_excel("data/SDR2025-data.xlsx", sheet = "Backdated SDG Index") %>% 
  filter(year >= 2000)

indicator_lookup_full <- indicator_lookup %>%
  left_join(sdg_table %>% 
              select(SDG, Title, Headline_Indicator_or_Proxy) %>%
              mutate(SDG = paste0("sdg", SDG)),
            by = c("goal" = "SDG")) %>%
  relocate(code, goal, short_name, label_guess, Title, Headline_Indicator_or_Proxy)

sdg_long <- sdg_raw %>%
  pivot_longer(
    cols = starts_with("sdg"),
    names_to  = "code",
    values_to = "value"
  ) %>%
  mutate(value = suppressWarnings(as.numeric(value)))  # in case of text like ".."
```



# 1) Country–year missingness 
```{r}
sdg_missing_cy <- sdg_long %>%
  group_by(Country, year) %>%
  summarise(
    total_indicators   = n(),
    missing_indicators = sum(is.na(value)),
    missing_rate       = missing_indicators / total_indicators,
    .groups = "drop"
  )

sdg_summary <- sdg_missing_cy %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing   = sd(missing_rate, na.rm = TRUE),
    min_missing  = min(missing_rate, na.rm = TRUE),
    max_missing  = max(missing_rate, na.rm = TRUE),
    n_countries  = n_distinct(Country),
    n_obs        = n()
  )
sdg_summary

kable(sdg_summary, caption = "Summary Statistics: SDG Data Missingness (2015–2024)")

```


**Interpretation** 
On average, `r round(sdg_summary$mean_missing * 100, 1)`% of SDG indicators 
are missing for a given country in a given year — meaning countries, on average,
report only about `r round((1 - sdg_summary$mean_missing) * 100, 1)` of 
SDG indicators.

The best-reported country-year had only `r round(sdg_summary$min_missing * 100, 1)`
% missing, while the worst had `r round(sdg_summary$max_missing * 100, 1)`%.

There’s moderate variation across countries and years — some countries are 
much more complete than others. The data has a standard deviation of 
`r round(sdg_summary$sd_missing * 100, 1)` percentage points.

The data includes `r round(n_countries = 208)` countries or territories, 
consistent with global SDG coverage.

#2: Variation across time 
```{r}
sdg_missing_cy %>%
  group_by(year) %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    median_missing = median(missing_rate, na.rm = TRUE)
  ) %>%
  ggplot(aes(x = year, y = mean_missing)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(title = "Average SDG Data Missingness Over Time",
       y = "Mean Missing Rate", x = "Year")
```

**Interpretation**
The Average SGD Data Missingness Over Time graph shows the mean proportion of 
missing Sustainable Development Goal (SDG) indicators per country-year.  
Between 2000 and 2018, the global average missingness steadily declined from 
roughly 60% to below 45%, reflecting improvements in data coverage and reporting
consistency under the early SDG framework.

However, after 2019, the trend reverses sharply.  
By 2025, the global mean missingness reaches `r round(max(missing_trend$mean_missing, na.rm = TRUE) * 100, 1)`%, 
indicating that only about `r round((1 - max(missing_trend$mean_missing, na.rm = TRUE)) * 100, 1)`
% of SDG indicators are being reported on average.  
This reversal coincides with post-pandemic disruptions, capacity shortfalls, 
and increasing political restrictions on data transparency.

Across the entire 2015–2024 period, countries report an average of 
`r round((1 - sdg_summary$mean_missing) * 100, 1)`% of SDG indicators, 
with a standard deviation of `r round(sdg_summary$sd_missing * 100, 1)` 
percentage points based on `r scales::comma(sdg_summary$n_obs)` 
country-year observations covering `r sdg_summary$n_countries` countries.  

Overall, the figure suggests that initial progress toward comprehensive SDG 
reporting  has been followed by a dramatic reversal, potentially linked to 
governance and transparency declines in the wake of global crises and democratic
backsliding.


# 3. Cross-Sectional Distribution
```{r}
sdg_missing_cy %>%
  filter(year == 2020) %>%
  ggplot(aes(x = missing_rate)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "white") +
  theme_bw() +
  labs(title = "Distribution of SDG Data Missingness (2020)",
       x = "Missing Rate", y = "Number of Countries")
```
       
**Interpretation**
The Distribution of SDG Data Missingness graph shows the distribution of SDG 
data missingness across countries in 2020.  
The average missingness rate that year was `r round(mean(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)`
%, with a standard deviation of `r round(sd(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)` percentage points.  

Most countries reported between `r round(min(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)`% and 
`r round(max(sdg_missing_cy$missing_rate[sdg_missing_cy$year == 2020], na.rm = TRUE) * 100, 1)`% of their SDG indicators, indicating moderate variation in reporting completeness 
across countries.  

The histogram shows a concentration around the mid-range (roughly 40–50%), 
suggesting that while many countries achieved partial reporting, 
full SDG data coverage remained uncommon in 2020.

#4. Between- vs Within-Country Variation
```{r}
sdg_missing_cy %>%
  group_by(Country) %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    n_years = n()
  ) %>%
  arrange(desc(mean_missing))
```


##Plotting a Few Cases
```{r}
top_countries <- c("Afghanistan", "Norway", "Kenya", "Brazil")

sdg_missing_cy %>%
  filter(Country %in% top_countries) %>%
  ggplot(aes(x = year, y = missing_rate, color = Country)) +
  geom_line(size = 1.2) +
  theme_bw() +
  labs(title = "Trends in SDG Missingness by Country",
       y = "Missing Rate", x = "Year")
```

# 5. Visualization: trend over time
## Average missingness over time
```{r}
missing_trend <- sdg_missing_cy %>%
  group_by(year) %>%
  summarise(
    mean_missing = mean(missing_rate, na.rm = TRUE),
    sd_missing = sd(missing_rate, na.rm = TRUE),
    n_countries = n_distinct(Country)
  )
```

## Line plot of the mean missing rate by year
```{r}
ggplot(missing_trend, aes(x = year, y = mean_missing)) +
  geom_line(color = "#2C3E50", size = 1.2) +
  geom_point(color = "#E74C3C", size = 2) +
  geom_ribbon(aes(ymin = mean_missing - sd_missing, ymax = mean_missing + sd_missing),
              fill = "skyblue", alpha = 0.2) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Average SDG Data Missingness Over Time (2000–2024)",
    x = "Year",
    y = "Mean Missing Rate",
    caption = "Source: Sustainable Development Report (2025)"
  )
```

**Interpretation** 
The Average SDG Data Missingness Over Time graph shows the trend in the 
global SDG data missingness rate from 2000 to 2025.  

The average share of missing indicators across countries is `r round(mean(missing_trend$mean_missing, na.rm = TRUE) * 100, 1)`%, 
with the lowest level around `r round(min(missing_trend$mean_missing, na.rm = TRUE) * 100, 1)`% in `r missing_trend$year[which.min(missing_trend$mean_missing)]` 
and the highest reaching `r round(max(missing_trend$mean_missing, na.rm = TRUE) * 100, 1)`% in `r missing_trend$year[which.max(missing_trend$mean_missing)]`.  

The steady decline in missingness through the mid-2010s suggests growing
statistical capacity and SDG data coverage, followed by a sharp reversal after 
2020—indicating a recent collapse in reporting consistency that may reflect 
post-pandemic disruptions and weakening transparency in some countries.
