---
title: "cleaning_output"
output: html_document
---

```{r}
# set working directory
setwd("~/Desktop/SIPA/Fall 25' - SIPA/Policy Data Analysis Using R/DataLoss_Proj")

# load packages
source("packages.R")
```

### Loading Data & Initial ID Checks
```{r}
# UNHASH ONLY IF INPUT DATA CHANGES [SET TO PROJECT DEFAULT DATA]
#source("wrangling/df_years2.0_Function_dataloss_proj.R") #load function
#merged_full <- df_years2.0(2015, 2023) #specify start & end year

# read merged dataframe
merged_full <- read_csv("data/output/merged_df.csv")

# UNIT TEST 1: Checking for missing identifier variables: country_code, country_name and year
check_missing_identifiers <- function(df){
  missing_id <- sum(is.na(df$country_code) | is.na(df$country_name) | is.na(df$year))
  print(paste("Number of missing identifiers:", missing_id))
  return(missing_id)}
check_missing_identifiers(merged_full)
```
**NOTE:** Default data loads a comprehensive inclusive merged dataframe from over 10 data sources. The default time-horizon for this dataframe spans 9 years (2015-2024). If input data changes, please un-hash the function source line and re-run to generate a new merged dataframe.

## 2) Selecting Relevant Variables
```{r}
# select relevant variables
merged_selected <- merged_full %>% 
  select(country_name, country_code, year, elect_dem, elect_dem_ert, lib_dem, part_dem, egal_dem, delib_dem, income_level, gdp_pc, log_gdppc, total_pop, log_pop, regime_type_2, regime_type_4, regime_type_10, regch_event, regch_genuine, dem_ep, dem_ep_pre_yr, dem_ep_start_yr, dem_ep_end_yr, aut_ep, aut_ep_pre_yr, aut_ep_start_yr, aut_ep_end_yr, electric_access, rural_pop_pct, rd_expenditure_pct, matches("^sdg\\d+")) %>% 
  arrange(country_code, year) %>% # arranging for easier viewing
  filter(year >= 2015 & year <= 2023) # filtering for years 2015-2023

# UNIT TEST 2: Checking for country_code and country_name inconsistencies
inconsistencies_u2 <- distinct(merged_selected, country_code, country_name) %>% 
  count(country_code) %>% 
  filter(n > 1)
print(inconsistencies_u2)
```
**NOTE:** This step narrows down the merged dataframe to only the variables relevant for analysis, including democracy indices, economic indicators, demographic data, regime types, and SDG indicators. It also includes a unit test to check for inconsistencies between country codes and country names, which is necessary for step 3.

## 3) Removing countries with no SDG data for any year (List A)
```{r}
# Defining SDG indicators to inspect 
sdg_indicators <- names(select(merged_selected, matches("^sdg\\d+")))

# Identifying countries with all SDG indicators missing for all years
all_missing_inds <- merged_selected %>% 
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(if_all(all_of(sdg_indicators), is.na))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()

# Exporting list for diagnostic purposes
write_csv(all_missing_inds, "wrangling/cleaning_output_diagnostics/A_missing_all_sdg_inds.csv")

# Removing countries these problematic countries from DF
cleaned_data <- merged_selected %>%
  anti_join(all_missing_inds, by = "country_code")
```
**NOTE:** This step first identifies then removes countries that have no SDG indicator data for any year in the dataset. This ensures that the final cleaned dataset only includes countries with at least some SDG data available. It also provides a diagnostic output(List A_...) listing the removed countries for transparency.

### UNIT TEST 3: Checking that no countries with all SDG indicators missing remain
```{r}
remaining_inconsistencies_u3 <- cleaned_data %>%
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(if_all(all_of(sdg_indicators), is.na))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
print(remaining_inconsistencies_u3) # should be empty
```
**NOTE:** This unit test verifies that no countries with all SDG indicators missing remain in the cleaned dataset after the removal step. If the output is empty, it confirms that the cleaning process was successful in this regard.

## 4) Removing Countries with no V-Dem Democracy Data for any year (List B)
```{r}
vdem_cols <- c('elect_dem', 'elect_dem_ert', 'lib_dem', 'part_dem', 'egal_dem', 'delib_dem')

# Identifying countries with all V-Dem democracy indicators missing for all years
all_missing_vdem <- cleaned_data %>%
  group_by(country_code, country_name) %>%
  summarise(
    all_missing = all(if_all(all_of(vdem_cols), is.na)),
    .groups = 'drop'
  ) %>%
  filter(all_missing) %>%
  select(country_code, country_name)

# Exporting list for diagnostic purposes
write_csv(all_missing_vdem, "wrangling/cleaning_output_diagnostics/B_missing_all_vdem_values.csv")

# Removing these problematic countries from cleaned_data
cleaned_data <- cleaned_data %>%
  filter(!country_code %in% all_missing_vdem$country_code)
```

### UNIT TEST 4: Checking that no countries with all V-Dem indicies missing remain
```{r}
remaining_inconsistencies_u4 <- cleaned_data %>%
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(if_all(all_of(vdem_cols), is.na))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
print(remaining_inconsistencies_u4) # should be empty
```

## 5) Missing Data Inspection (Main Xs + Controls)
***Specific Main Xs:*** elect_dem, lib_dem, part_dem, egal_dem, delib_dem

***Specific Controls:*** gdp_pc, total_pop, electric_access, rural_pop_pct, rd_expenditure_pct, regime_type_2, regime_type_4, income_level
```{r}
# Checking missingness across panel
all_country_years <- expand.grid(
  country_code = unique(cleaned_data$country_code),
  year = unique(cleaned_data$year)
)
check_df <- left_join(all_country_years, cleaned_data, 
                      by = c("country_code", "year"))

# Change variable name as needed
inspect_x <- "elect_dem" # Example: Check for missing values in 'gdp_pc'

missing_varX <- check_df %>% 
  arrange(country_code, year) %>%
  select(country_code, year, all_of(inspect_x)) %>%
  filter(is.na(.data[[inspect_x]]))
print(missing_varX)
```

## 4.2) Missing Data Inspection (Main Xs + Controls) - Pivot Table 
```{r}
# Defining control variables
controls <- c("elect_dem", "lib_dem", "part_dem", "egal_dem", "delib_dem", "gdp_pc", "total_pop", "electric_access", "rural_pop_pct", "rd_expenditure_pct", "regime_type_2", "regime_type_4", "income_level")

# Df with control variables
df_controls <- cleaned_data %>% select(country_code, country_name, year, all_of(controls))

# Pivot table of NA counts for each control variable by country
na_counts_controls <- df_controls %>%
  group_by(country_code, country_name) %>%
  summarise(
    across(all_of(controls), ~sum(!is.na(.)), .names = "{.col}")
  )

View(na_counts_controls)
```


## 5) Data Types and Final Checks
```{r}
# converting data types for panel suitability 
cleaned_data$country_code <- as.character(cleaned_data$country_code)
cleaned_data$year <- as.integer(cleaned_data$year)

# variables to factor (categorical/dummy)
cleaned_data <- cleaned_data %>%
  mutate( 
    # All regime and episode variables factorized
    regime_type_2 = as.factor(regime_type_2),
    regime_type_4 = as.factor(regime_type_4),
    regime_type_10 = as.factor(regime_type_10),
    regch_event = as.factor(regch_event),
    regch_genuine = as.factor(regch_genuine),
    dem_ep = as.factor(dem_ep),
    dem_ep_pre_yr = as.factor(dem_ep_pre_yr),
    dem_ep_start_yr = as.factor(dem_ep_start_yr),
    dem_ep_end_yr = as.factor(dem_ep_end_yr),
    aut_ep = as.factor(aut_ep),
    aut_ep_pre_yr = as.factor(aut_ep_pre_yr),
    aut_ep_start_yr = as.factor(aut_ep_start_yr),
    aut_ep_end_yr = as.factor(aut_ep_end_yr)
  )
```

## 6) Uploading Missingness Variable 
```{r}
# source missingness variable creation script
source('___missingness_var.R-script____')

## [INSERT MERGING CODE HERE TO MERGE MISSINGNESS VARS INTO cleaned_data]

```


## 6) Exporting Cleaned Data
```{r}
# export cleaned data
write_csv(cleaned_data, "data/output/cleaned_data.csv")
```


