
---
title: "Democratic Backsliding and SDG Data Missingness"
subtitle: "Preliminary descriptive & TWFE results"
author: "Priskila Nandita, Naz Hussein, Sevastian Sanchez"
institute: "Columbia University | SIPA"
output:
  xaringan::moon_reader:
    self_contained: true
    lib_dir: libs
    nature:
      mathjax: "default"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---   
<style>
.underline {
  text-decoration: underline;
}
</style>

<style>
.remark-slide-content:not(.title-slide) {
  background-color: #F0F2E2; /* Light lavender pastel */
}
</style>

<style>
.title-slide {
  background-image:
    linear-gradient(rgba(0,0,0,0.4), rgba(0,0,0,0.4)),
    url("https://images.unsplash.com/photo-1584931423298-c576fda54bd2?q=80&w=1740&auto=format&fit=crop&ixlib=rb-4.1.0&ixid=M3wxMjA3fDB8MHxwaG90by1wYWdlfHx8fGVufDB8fHx8fA%3D%3D");
  background-size: cover;
  background-position: center;
}
</style>

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  dev = "png",
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(fixest)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(knitr)
library(broom)
library(Cairo) 
library(modelsummary)
library(stargazer)
library(kableExtra)
library(lmtest)
library(multiwayvcov)
library(sandwich)

# Load final combined panel data (must be in same folder as this Rmd)
load("sdg_democracy_paneldata.RData")

# Assume main panel is called df2
stopifnot(exists("df2"))

# Basic metadata
n_countries <- dplyr::n_distinct(df2$country_code)
years_min   <- min(df2$year, na.rm = TRUE)
years_max   <- max(df2$year, na.rm = TRUE)

# Helper: build goal-level missingness long-form from sdg* columns
sdg_vars <- grep("^sdg", names(df2), value = TRUE)

goal_missing <- function(g) {
  goal_cols <- grep(paste0("^sdg", g, "_"), names(df2), value = TRUE)
  if (length(goal_cols) == 0) return(NULL)
  df2 %>%
    mutate(
      goal = paste0("SDG", g),
      n_ind = length(goal_cols),
      n_miss = apply(select(., all_of(goal_cols)), 1,
                     function(x) sum(is.na(x))),
      prop_missing = n_miss / n_ind
    ) %>%
    select(
      country_code,
      country_name,
      year,
      income_level,
      goal,
      n_ind,
      n_miss,
      prop_missing,
      elect_dem
    )
}

goal_miss_df <- dplyr::bind_rows(lapply(1:17, goal_missing))
```

#Country in Autocratization Episodes Over Time

<div style="text-align:center;">


```{r}
#install.packages("plotly")
library(plotly)
```

.pull-center[
```{r aut_map, fig.align='center', fig.height=6}

auto_map <- df2 %>%
  group_by(country_code, year) %>%
  summarise(
    aut_status = max(aut_ep, na.rm = TRUE),  # 1 = autocratized, 0 = not
    .groups = "drop"
  )

p <- plot_ly(
  data = auto_map,
  type = "choropleth",
  locations = ~country_code,      # ISO3 codes
  locationmode = "ISO-3",
  z = ~aut_status,                # 0 or 1
  frame = ~year,                  # <- this creates the year slider
  colorscale = list(
    c(0,  "#f0f0f0"),             # light grey = not autocratized
    c(1,  "#e41a1c")              # red = in autocratization episode
  ),
  showscale = FALSE,
  hovertemplate = paste(
    "Country: %{location}<br>",
    "Year: %{frame}<br>",
    "Autocratization: %{z}<extra></extra>"
  )
) %>%
  layout(
    title = "Countries in Autocratization Episodes Over Time",
    geo = list(
      showframe = FALSE,
      showcoastlines = FALSE,
      projection = list(type = "robinson")
    )
  )

p
```
]

---
#Average SDG missingness over time by democratic trajectory

```{r processing}
# 1. Classify each country by episode history
dem_groups <- df2 %>%
  group_by(country_code, country_name) %>%
  summarise(
    ever_dem = any(dem_ep == 1, na.rm = TRUE),
    ever_aut = any(aut_ep == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    trajectory = dplyr::case_when(
      ever_aut & !ever_dem ~ "Backsliding only",
      ever_dem & !ever_aut ~ "Democratizing only",
      ever_dem & ever_aut  ~ "Backsliding & democratizing",
      TRUE                 ~ "No major episode"
    ),
    trajectory = factor(
      trajectory,
      levels = c("Backsliding only",
                 "Democratizing only",
                 "Backsliding & democratizing",
                 "No major episode")
    )
  )

# 2. Attach trajectory to panel data
df_traj <- df2 %>%
  left_join(dem_groups, by = c("country_code", "country_name"))
```

.pull-center[
```{r, fig.align='center', out.width="80%", fig.height=4}
avg_missing_simple <- df_traj %>%
  mutate(
    treated_group = if_else(
      trajectory %in% c("Backsliding only", "Backsliding & democratizing"),
      "Ever autocratized",
      "Never autocratized"
    )
  ) %>%
  group_by(year, treated_group) %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    .groups = "drop"
  )


ggplot(avg_missing_simple,
       aes(x = year, y = mean_missing, color = treated_group)) +
  geom_line(size = 1) +
  geom_point() +
  labs(
    title = "Average SDG Missingness: Ever vs Never Autocratized",
    x = "Year",
    y = "Mean proportion of SDG indicators missing",
    color = "Group"
  ) +
  theme_minimal()
```
]

---
# Staggered DiD: PRF


We estimate:

$$Y_{it} = \alpha_i + \lambda_t + ATT_{g,t} \cdot D_{it} + X_{it}\gamma + \varepsilon_{it}$$

Where:

- $Y_{it}$: SDG missingness in country $i$, year $t$ 
- $D_{it}$: indicator that country \(i\) is treated (autocratized) in year \(t\)
- $ATT_{g,t}$: group-time average treatment effect for countries first treated at time \(g\)
- $X_{it}$: socioeconomic controls (GDPpc, electricity access, rural pop share, R&D spending)
- $\alpha_i$: country fixed effects
- $\lambda_t$: year fixed effects
- $\varepsilon_{it}$: error term

This formulation allows treatment effects to differ depending on the year  
each country autocratized.

---
#Future Modeling Strategy

To further strengthen causal identification, we plan to:

- Restrict sample to "sufficiently democratic" countries pre-episode to avoid bias from very low-capacity/borderline states.
- Heterogeneous treatment effects
Allow effect sizes to differ by: baseline democracy level, statistical capacity, region, income group  
- Futher analysis.
We will complement the quantitative results with a qualitative  event analysis of several illustrative country cases that  experienced autocratization (e.g., Hungary, Turkey, India).

---

