---
title: "SDG Missingness, Democracy, and Transparency"
author: "Priskila Nandita"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)

library(tidyverse)
library(readr)
library(tidyr)
library(ggplot2)
library(fixest)
library(viridis)
library(dplyr)
library(Hmisc)
```

# Load Data

```{r}
df <- read_csv("cleaned_data.csv")

sdg_cols <- df %>%
  select(starts_with("sdg")) %>%
  colnames()

length(sdg_cols)
```

# Construct SDG Missingness Dataset

```{r}
df <- df %>%
  mutate(
    n_sdg_indicators = length(sdg_cols),
    n_sdg_missing    = apply(select(., all_of(sdg_cols)), 1, function(x) sum(is.na(x))),
    prop_sdg_missing = n_sdg_missing / n_sdg_indicators
  )
```

## Overall distribution of missingness

```{r}
df %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    sd_missing   = sd(prop_sdg_missing, na.rm = TRUE),
    min_missing  = min(prop_sdg_missing, na.rm = TRUE),
    max_missing  = max(prop_sdg_missing, na.rm = TRUE)
  )

# Histogram
ggplot(df, aes(x = prop_sdg_missing)) +
  geom_histogram(bins = 30) +
  labs(
    x = "Proportion of SDG indicators missing",
    y = "Number of country-years",
    title = "Distribution of SDG data missingness"
  )
```

## Missingness by SDG goal
We compute proportion missing per goal (1–17) for each country-year.

```{r}
# Construct per-goal missingness for each country-year

sdg_cols <- grep("^sdg", names(df2), value = TRUE)

for (g in 1:17) {
  # all indicator columns for this goal in df2
  goal_cols <- grep(paste0("^sdg", g, "_"), names(df2), value = TRUE)
  
  # skip if this goal has no indicators in your data
  if (length(goal_cols) == 0) next
  
  # count how many indicators under this goal are missing in each row
  df2[[paste0("n_miss_SDG", g)]] <- apply(df2[goal_cols], 1, function(x) sum(is.na(x)))
}

for (g in 1:17) {
  goal_cols <- grep(paste0("^sdg", g, "_"), names(df2), value = TRUE)
  if (length(goal_cols) == 0) next
  
  # proportion of indicators missing for that goal
  df2[[paste0("prop_miss_SDG", g)]] <-
    df2[[paste0("n_miss_SDG", g)]] / length(goal_cols)
}
```

```{r}

goal_missing <- function(g) {
  goal_cols <- grep(paste0("^sdg", g, "_"), names(df), value = TRUE)
  df %>%
    mutate(
      goal = paste0("SDG", g),
      n_goal_ind = length(goal_cols),
      n_goal_missing = apply(select(., all_of(goal_cols)), 1, function(x) sum(is.na(x))),
      prop_missing = n_goal_missing / n_goal_ind
    ) %>%
    select(country_code, country_name, year, income_level, goal, prop_missing)
}

goal_miss_df <- bind_rows(lapply(1:17, goal_missing))

# Average missingness per goal across all country-years
goal_summary <- goal_miss_df %>%
  group_by(goal) %>%
  summarise(
    mean_prop_missing = mean(prop_missing, na.rm = TRUE),
    sd_prop_missing   = sd(prop_missing, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_prop_missing))

goal_summary

#Bar plot of missingness by SDG goal

ggplot(goal_summary, aes(x = reorder(goal, mean_prop_missing), y = mean_prop_missing)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Average Proportion of Missing SDG Indicators, by Goal",
    x = "SDG Goal",
    y = "Mean Missingness"
  )
```

```{r}
#Pattern across country-income groups

#Overall missingness
income_missing_summary <- df %>%
  group_by(income_level) %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    sd_missing   = sd(prop_sdg_missing, na.rm = TRUE),
    median_missing = median(prop_sdg_missing, na.rm = TRUE),
    n_obs = n()
  ) %>%
  arrange(desc(mean_missing))

income_missing_summary


#Bar Plot with 95% CI
income_ci <- df %>%
  group_by(income_level) %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    se_missing   = sd(prop_sdg_missing, na.rm = TRUE) / sqrt(n()),
    lower_ci     = mean_missing - 1.96 * se_missing,
    upper_ci     = mean_missing + 1.96 * se_missing
  )

ggplot(income_ci, aes(x = income_level, y = mean_missing)) +
  geom_col(fill = "steelblue") +
  geom_errorbar(
    aes(ymin = lower_ci, ymax = upper_ci),
    width = 0.2
  ) +
  labs(
    title = "Average SDG Missingness by Income Group (95% CI)",
    x = "Income Group",
    y = "Mean proportion missing"
  ) +
  theme_minimal()

```
## Overall Average Missingness over time 
```{r}
overall_missing_over_time <- df %>%
  group_by(year) %>%
  summarise(
    mean_missing = mean(prop_sdg_missing, na.rm = TRUE),
    sd_missing = sd(prop_sdg_missing, na.rm = TRUE),
    n = n()
  )


ggplot(overall_missing_over_time, aes(x = year, y = mean_missing)) +
  geom_line(color = "steelblue", size = 1.2) +
  geom_point(color = "steelblue") +
  labs(
    title = "Average SDG Missingness Over Time",
    x = "Year",
    y = "Mean Proportion Missing"
  ) +
  theme_minimal()

overall_missing_over_time

```

#Goal-Level Average Missingness Over Time

```{r}
goal_missing_over_time <- goal_miss_df %>%
  group_by(goal, year) %>%
  summarise(
    mean_missing = mean(prop_missing, na.rm = TRUE),
    sd_missing = sd(prop_missing, na.rm = TRUE),
    n = n()
  ) %>%
  ungroup()

goal_missing_over_time

ggplot(goal_missing_over_time,
       aes(x = year, y = mean_missing)) +
  geom_line(color = "darkred") +
  geom_point(color = "darkred", size = 0.8) +
  facet_wrap(~ goal, scales = "free_y") +
  labs(
    title = "Average SDG Missingness Over Time by Goal",
    x = "Year",
    y = "Mean Proportion Missing"
  ) +
  theme_minimal()

```

```{r}
#Patterns across per goal missingness by income group

goal_missing <- function(g) {
  goal_cols <- grep(paste0("^sdg", g, "_"), names(df), value = TRUE)
  
  df %>%
    mutate(
      goal = paste0("SDG", g),
      n_ind = length(goal_cols),
      n_miss = apply(select(., all_of(goal_cols)), 1, function(x) sum(is.na(x))),
      prop_missing = n_miss / n_ind
    ) %>%
    select(
      country_code,
      country_name,
      year,
      income_level,   # <-- ADD THIS
      goal,
      prop_missing
    )
}

goal_miss_df <- dplyr::bind_rows(lapply(1:17, goal_missing))

goal_missing_over_time <- goal_miss_df %>%
  dplyr::group_by(goal, year) %>%
  dplyr::summarise(
    mean_missing = mean(prop_missing, na.rm = TRUE),
    sd_missing   = sd(prop_missing, na.rm = TRUE),
    n = dplyr::n()
  ) %>%
  dplyr::ungroup()

```

```{r}
goal_income_summary <- goal_miss_df %>%
  group_by(goal, income_level) %>%
  summarise(mean_missing = mean(prop_missing, na.rm = TRUE))

goal_income_summary
```

```{r}
ggplot(goal_income_summary,
       aes(x = income_level, y = mean_missing, fill = income_level)) +
  geom_col() +
  facet_wrap(~ goal, ncol = 4) +
  labs(
    title = "Average SDG Missingness by Goal and Income Group",
    x = "Income Group",
    y = "Proportion Missing"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 10, face = "bold")
  )
```
# Demcoractic variables

## World map of democracy levels

```{r}
#install.packages(c("sf", "rnaturalearth", "rnaturalearthdata", "ggplot2", "dplyr"))

library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

world <- ne_countries(scale = "medium", returnclass = "sf")


df_latest <- df %>%
  filter(year == max(year)) %>%
  select(country_code, elect_dem)

world_dem <- world %>%
  left_join(df_latest, by = c("iso_a3" = "country_code"))

ggplot(data = world_dem) +
  geom_sf(aes(fill = elect_dem), color = "gray30", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", na.value = "lightgray") +
  theme_minimal() +
  labs(
    title = "World Map of Electoral Democracy (Latest Year)",
    fill  = "Electoral Democracy"
  ) +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "right"
  )

```

## Gobal democracy trend over time

```{r}
df %>%
  group_by(year) %>%
  summarise(global_mean = mean(elect_dem, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = global_mean)) +
  geom_line(size = 1, color = "darkred") +
  geom_point() +
  theme_minimal() +
  labs(
    title = "Global Trend in Electoral Democracy Over Time",
    x = "Year",
    y = "Global mean elect_dem"
  )
```

## How many countries experienced democratization / backsliding?

```{r}
dem_groups <- df %>%
  group_by(country_code, country_name) %>%
  summarise(
    ever_dem = any(dem_ep == 1, na.rm = TRUE),
    ever_aut = any(aut_ep == 1, na.rm = TRUE)
  ) %>%
  mutate(
    trajectory = case_when(
      ever_aut & !ever_dem ~ "Backsliding only",
      ever_dem & !ever_aut ~ "Democratizing only",
      ever_dem & ever_aut  ~ "Both episodes",
      TRUE                 ~ "No major episode"
    )
  )


table(dem_groups$trajectory)


ggplot(dem_groups, aes(x = trajectory, fill = trajectory)) +
  geom_bar() +
  labs(
    title = "Number of Countries by Democratic Trajectory (V-Dem Episodes)",
    x = "",
    y = "Number of Countries"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

only_backsliding <- dem_groups %>%
  filter(trajectory == "Backsliding only") %>%
  arrange(country_name)

only_backsliding
```

## Map autocratization episodes

```{r}
df_auto <- df %>%
  group_by(country_code) %>%
  summarise(ever_aut = any(aut_ep == 1, na.rm = TRUE))

world_auto <- world %>%
  left_join(df_auto, by = c("iso_a3" = "country_code"))

ggplot(world_auto) +
  geom_sf(aes(fill = ever_aut)) +
  scale_fill_manual(values = c("lightgray", "red")) +
  theme_minimal() +
  labs(title = "Countries That Experienced Autocratization Episodes")

```

# Missingness vs Democracy

## Correlation and scatter plots: democracy vs missingness

```{r}
df %>%
  select(prop_sdg_missing, elect_dem, lib_dem, part_dem, egal_dem, delib_dem) %>%
  cor(use = "pairwise.complete.obs")

ggplot(df, aes(x = elect_dem, y = prop_sdg_missing)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Electoral democracy index (V-Dem)",
    y = "Proportion of SDG indicators missing",
    title = "Relationship between democracy and SDG data missingness"
  )

```

## Other plots for democratic variables

# 6. Two-Way Fixed Effects Models with staggered treatment

Question: Does entering an autocratization episode increase SDG data missingness?

## Define treatment timing (staggered adoption)
We use aut_ep_start_yr as the group (treatment year). For countries that never backslide, this is NA.

```{r}
treat_df <- df %>%
  group_by(country_code) %>%
  summarise(
    G_aut = suppressWarnings(min(aut_ep_start_yr[aut_ep == 1], na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  mutate(
    G_aut = ifelse(is.infinite(G_aut), NA, G_aut)   # if no episode, set to NA
  )

df2 <- df %>%
  left_join(treat_df, by = "country_code") %>%
  mutate(
    # TWFE-like treatment indicator: treated once year >= G_aut
    D_aut = ifelse(!is.na(G_aut) & year >= G_aut, 1, 0)
  )
```

## TFWE Model 
Outcome: prop_sdg_missing
Treatment: D_aut
Fixed effects: country and year
Errors: clustered by country

```{r}
twfe_model <- feols(
  prop_sdg_missing ~ D_aut | country_code + year,
  cluster = ~ country_code,
  data = df2
)

summary(twfe_model)

twfe_model2 <- feols(
  prop_sdg_missing ~ D_aut + log_gdppc | country_code + year,
  cluster = ~ country_code,
  data = df2
)
summary(twfe_model2)

twfe_model3 <- feols(
  prop_sdg_missing ~ D_aut + log_gdppc + electric_access + rural_pop_pct + rd_expenditure_pct | country_code + year,
  cluster = ~ country_code,
  data = df2
)
summary(twfe_model3)


```
PRF

```{r}
\[
Y_{it} = \alpha_i + \lambda_t 
       + \beta D_{it} 
       + \gamma_1 \log(\text{GDPpc}_{it})
       + \gamma_2 \text{electric\_access}_{it}
       + \gamma_3 \text{rural\_pop\_pct}_{it}
       + \gamma_4 \text{rd\_expenditure\_pct}_{it}
       + \varepsilon_{it},
\]
```



After a country enters an autocratization episode, the proportion of SDG indicators missing increases by about 0.94 percentage points, relative to:
- its own pre-autocratization levels, and
- other countries that never autocratized or have not yet autocratized.

The effect is small, and the p-value (~0.057) means:
- borderline significance at the 10% level,
- not statistically significant at the 5% level.

So: autocratization may increase SDG missingness slightly, but evidence is not strong.


```{r}
save(df2, file = "sdg_democracy_paneldata.RData")
```
























