---
title: "Callaway & Sant'Anna DiD — SDG Data Missingness & Democratic Backsliding"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r packages}
library(tidyverse)
library(did)
library(ggplot2)
library(patchwork)
```

---

# 1. Load Data

```{r load-data}
# panel <- read_csv("data/output/panel_es_data.csv")
source("causal_inference_mods/causal_inf_data_prep.R")


# number of NAs for each control variable using colSums and is.na
colSums(is.na(panel %>% select(elect_dem, lib_dem, spi_overall, log_gdppc,
                               log_pop, luminosity, electric_access, rural_pop_pct)))
```

---

# 2. Build CS Panel

Key decisions encoded here:

- **Left-censored countries** (autocratization episode starts before 2015) are excluded
  because we cannot observe their pre-treatment trajectory in this panel.
- **`gname`** = first autocratization start year (`aut_cohort`), or 0 for never-treated.
  This is held constant across all rows for each country (CS requirement).
- **Cohort restriction**: only cohorts 2016–2020 are retained as treated units,
  plus all never-treated (gname == 0). Cohorts 2015 (treated at panel start, no
  pre-period), and 2021–2023 (too few treated units and almost no post-period)
  are excluded as they cannot be identified in the CS framework.
- **`idname`** = `country_id` (numeric).
- **`tname`** = `year` (integer).

```{r build-panel}
# # Step 1: Flag left-censored countries
# panel <- panel %>%
#   mutate(
#     left_censored = (!is.na(aut_start_filled) & aut_start_filled < 2015)
#   )
# 
# cat("Left-censored countries (episode starts before 2015):",
#     n_distinct(panel$country_code[panel$left_censored]), "\n")
# 
# # Step 2: Build CS panel
# panel_cs <- panel %>%
#   filter(!left_censored) %>%
#   mutate(year = as.integer(year)) %>%
#   rename(idname = country_id,
#          tname  = year)
# 
# # Step 3: Assign constant gname per country (first episode start year; 0 = never-treated)
# g_by_country <- panel_cs %>%
#   group_by(idname) %>%
#   summarise(
#     gname = ifelse(
#       any(aut_cohort > 0, na.rm = TRUE),
#       min(aut_cohort[aut_cohort > 0], na.rm = TRUE),
#       0L
#     ),
#     .groups = "drop"
#   )
# 
# panel_cs <- panel_cs %>%
#   select(-any_of("gname")) %>%
#   left_join(g_by_country, by = "idname")
# 
# # Step 4: Restrict to identifiable cohorts only
# panel_cs <- panel_cs %>%
#   filter(
#     gname == 0 | gname %in% 2016:2020
#   )
```

```{r cohort-check}
# # Verify gname is constant per country
# n_multi_g <- panel_cs %>%
#   distinct(idname, gname) %>%
#   count(idname) %>%
#   filter(n > 1) %>%
#   nrow()
# 
# cat("Countries with inconsistent gname (should be 0):", n_multi_g, "\n")
# 
# # Cohort distribution
# panel_cs %>%
#   distinct(idname, gname) %>%
#   count(gname, name = "n_countries") %>%
#   arrange(gname) %>%
#   print()
```

---

# 3. SPI Moderator Variable

```{r spi-moderator}
# # Compute baseline SPI per country (first non-missing observation)
# spi_baseline <- panel_cs %>%
#   arrange(idname, tname) %>%
#   group_by(idname) %>%
#   summarise(spi_base = first(na.omit(spi_overall)), .groups = "drop")
# 
# # Median-split into High/Low SPI
# spi_median <- median(spi_baseline$spi_base, na.rm = TRUE)
# 
# panel_cs <- panel_cs %>%
#   left_join(spi_baseline, by = "idname") %>%
#   mutate(
#     spi_hilow = case_when(
#       spi_base >= spi_median ~ "High SPI",
#       spi_base <  spi_median ~ "Low SPI",
#       TRUE ~ NA_character_
#     )
#   )
# 
# cat("SPI median (baseline):", round(spi_median, 2), "\n")
```

---

# 4. Outcome Variables

```{r outcomes}
# # Overall outcome
# outcome_overall <- "prop_sdg_missing"
# 
# # Goal-level outcomes
# sdg_goals <- paste0("prop_miss_SDG", 1:17)
```

---

# 5. Overall SDG Missingness — CS Estimation

> **Note on `aggte()`**: Due to a known internal issue in some versions of the
> `did` package where `aggte()` triggers `Error in get(gname) : invalid first
> argument`, aggregation is performed manually from the group-time ATT object
> rather than via `aggte()`. Results are substantively identical.

## 5a. Run `att_gt()`

```{r cs-overall, cache=TRUE}
cs_overall <- att_gt(
  yname = "prop_sdg_missing",
  tname = "tname", # time variable (year)
  idname = "idname", # ID tells which rows belong to the same country over time
  gname = "gname", # ID of first year it is treated (cohort)
  xformla = ~ log_pop + electric_access + log_luminosity + rural_pop_pct,
  control_group = "nevertreated", # control group = countries never having aut_ep
  anticipation = 0, # treatment effect = start-year onset (no anticipation)
  data = panel_cs %>% filter(!is.na(prop_sdg_missing)) %>% as.data.frame(),
  est_method = "dr",
  clustervars = "idname",
  print_details = TRUE
)

summary(cs_overall)
```

## 5b. Manual Overall ATT (post-treatment average)

```{r overall-att-manual}
gt_overall <- data.frame(
  group = cs_overall$group,
  time = cs_overall$t,
  att = cs_overall$att,
  se = cs_overall$se
) %>%
  mutate(event_time = time - group)   # event time

# Simple average ATT across all post-treatment (g,t) pairs
overall_att_simple <- gt_overall %>%
  filter(event_time >= 0) %>%
  summarise(
    mean_att = mean(att, na.rm = TRUE),
    mean_se = mean(se, na.rm = TRUE)
  )

cat("Simple overall ATT (event_time >= 0):\n")
print(overall_att_simple)
```

## 5c. Dynamic ATT by Event Time

```{r dynamic-att-manual}
# Average ATT(g,t) by event time event_time (weighted equally across cohorts)
dynamic_overall <- gt_overall %>%
  group_by(event_time) %>%
  summarise(
    att_e = mean(att, na.rm = TRUE),
    se_e = sqrt(mean(se^2, na.rm = TRUE)),
    n_gt = n(),
    .groups = "drop"
  ) %>%
  filter(event_time >= -4, event_time <= 5) %>%
  mutate(
    ci_lo = att_e - 1.96 * se_e,
    ci_hi = att_e + 1.96 * se_e
  )

print(dynamic_overall)
```

## 5d. Event-Study Plot — Overall SDG Missingness

```{r plot-overall-dynamic, fig.width=9, fig.height=5}
ggplot(dynamic_overall, aes(x = event_time, y = att_e)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = -0.5, linetype = "dashed", colour = "red", alpha = 0.6) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#4393c3", alpha = 0.2) +
  geom_line(colour = "#2166ac", linewidth = 0.8) +
  geom_point(colour = "#2166ac", size = 2.5) +
  scale_x_continuous(breaks = -4:5) +
  labs(
    title    = "CS Event-Study: Overall SDG Missingness",
    subtitle = "Doubly-robust DiD | Never-treated controls | Anticipation = 1\nShaded band = 95% CI | Red dashed line = treatment onset",
    x        = "Years since start of autocratization episode",
    y        = "Average ATT(g,t): Change in prop_sdg_missing"
  ) +
  theme_minimal(base_size = 13)
```

## 5e. Group-Level ATT (by Treatment Cohort)

```{r group-att-manual}
group_overall <- gt_overall %>%
  filter(event_time >= 0) %>%
  group_by(group) %>%
  summarise(
    att_group = mean(att, na.rm = TRUE),
    se_group  = sqrt(mean(se^2, na.rm = TRUE)),
    n_gt      = n(),
    .groups   = "drop"
  ) %>%
  mutate(
    ci_lo = att_group - 1.96 * se_group,
    ci_hi = att_group + 1.96 * se_group
  )

print(group_overall)
```


```{r plot-group-att, fig.width=8, fig.height=4}
ggplot(group_overall, aes(x = factor(group), y = att_group)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_errorbar(aes(ymin = ci_lo, ymax = ci_hi), width = 0.15, colour = "#2166ac") +
  geom_point(colour = "#2166ac", size = 3) +
  labs(
    title = "CS Group-Level ATT: Overall SDG Missingness",
    subtitle = "Average post-treatment ATT by first-episode cohort",
    x = "Treatment Cohort (first episode start year)",
    y = "ATT"
  ) +
  theme_minimal(base_size = 13)
```

---

# 6. Heterogeneous Effects by Income Level

```{r income-loop, cache=TRUE}
income_levels <- c("L", "LM", "UM", "H")

cs_income_results <- list()

for (inc in income_levels) {
  df_inc <- panel_cs %>%
    filter(income_level == inc, !is.na(prop_sdg_missing)) %>%
    as.data.frame()

  n_treated <- df_inc %>% distinct(idname, gname) %>% filter(gname > 0) %>% nrow()

  if (n_treated < 3) {
    message("Skipping income group ", inc, ": only ", n_treated, " treated countries.")
    next
  }

  cs_inc <- tryCatch(
    att_gt(
      yname = "prop_sdg_missing",
      tname = "tname",
      idname = "idname",
      gname = "gname",
      xformla = ~ log_pop + electric_access + log_luminosity + rural_pop_pct,
      #log_pop + electric_access + log_luminosity + rural_pop_pct
      control_group = "nevertreated",
      anticipation = 1,
      data = df_inc,
      est_method = "dr",
      clustervars = "idname",
      print_details = FALSE
    ),
    error = function(event_time) {
      message("Income group ", inc, " failed: ", event_time$message)
      NULL
    }
  )

  if (!is.null(cs_inc)) {
    cs_income_results[[inc]] <- cs_inc
  }
}

cat("Income groups successfully estimated:", paste(names(cs_income_results), collapse = ", "), "\n")
```

```{r income-dynamic-tables}
# Build dynamic tables for each income group
income_dynamic <- imap_dfr(cs_income_results, function(cs_obj, inc) {
  data.frame(
    income  = inc,
    group   = cs_obj$group,
    time    = cs_obj$t,
    att     = cs_obj$att,
    se      = cs_obj$se
  ) %>%
    mutate(event_time = time - group) %>%
    filter(event_time >= -4, event_time <= 5) %>%
    group_by(income, event_time) %>%
    summarise(
      att_e = mean(att, na.rm = TRUE),
      se_e  = sqrt(mean(se^2, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    mutate(
      ci_lo = att_e - 1.96 * se_e,
      ci_hi = att_e + 1.96 * se_e
    )
})

print(income_dynamic)
```

```{r income-dynamic-plots, fig.width=11, fig.height=7}
ggplot(income_dynamic, aes(x = event_time, y = att_e, colour = income, fill = income)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = -1, linetype = "dashed", colour = "red", alpha = 0.5) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = -4:5) +
  facet_wrap(~ income, ncol = 2) +
  labs(
    title    = "CS Event-Study by Income Level: Overall SDG Missingness",
    subtitle = "Doubly-robust DiD | Never-treated controls | Anticipation = 1",
    x        = "Years since autocratization onset",
    y        = "ATT"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

---

# 7. Goal-Level Models

```{r goal-loop, cache=TRUE}
cs_goal_results <- list()

for (g in sdg_goals) {
  df_g <- panel_cs %>%
    filter(!is.na(.data[[g]])) %>%
    as.data.frame()

  n_treated <- df_g %>% distinct(idname, gname) %>% filter(gname > 0) %>% nrow()

  if (n_treated < 3) {
    message("Skipping ", g, ": only ", n_treated, " treated countries with non-missing outcome.")
    next
  }

  cs_g <- tryCatch(
    att_gt(
      yname         = g,
      tname         = "tname",
      idname        = "idname",
      gname         = "gname",
      xformla       = ~ log_pop + electric_access + log_luminosity + rural_pop_pct,
      control_group = "nevertreated",
      anticipation  = 1,
      data          = df_g,
      est_method    = "dr",
      clustervars   = "idname",
      print_details = FALSE
    ),
    error = function(event_time) {
      message(g, " failed: ", event_time$message)
      NULL
    }
  )

  if (!is.null(cs_g)) {
    cs_goal_results[[g]] <- cs_g
  }
}

cat("Goals successfully estimated:", length(cs_goal_results), "/", length(sdg_goals), "\n")
```

## 7a. Tidy dynamic ATT table by goal

```{r goal-dynamic-table}
goal_dynamic_df <- imap_dfr(cs_goal_results, function(cs_obj, g) {
  data.frame(
    goal  = g,
    group = cs_obj$group,
    time  = cs_obj$t,
    att   = cs_obj$att,
    se    = cs_obj$se
  ) %>%
    mutate(
      event_time        = time - group,
      goal_num = as.integer(str_extract(g, "[0-9]+"))
    ) %>%
    filter(event_time >= -4, event_time <= 5) %>%
    group_by(goal, goal_num, event_time) %>%
    summarise(
      att_e = mean(att, na.rm = TRUE),
      se_e  = sqrt(mean(se^2, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    mutate(
      ci_lo = att_e - 1.96 * se_e,
      ci_hi = att_e + 1.96 * se_e
    )
})

# ATT at onset (event_time == 0) by goal
goal_dynamic_df %>%
  filter(event_time == 0) %>%
  arrange(desc(att_e)) %>%
  print(n = 20)
```

## 7b. Heatmap of post-treatment ATTs by SDG Goal

```{r goal-heatmap, fig.width=11, fig.height=6}
goal_dynamic_df %>%
  filter(event_time >= 0, event_time <= 5) %>%
  mutate(
    sig      = (ci_lo > 0 | ci_hi < 0),
    sig_mark = if_else(sig, "*", "")
  ) %>%
  ggplot(aes(x = event_time, y = reorder(goal, -goal_num), fill = att_e)) +
  geom_tile(colour = "white") +
  geom_text(aes(label = sig_mark), colour = "black", size = 4) +
  scale_fill_gradient2(
    low  = "#2166ac", mid = "white", high = "#d73027",
    name = "ATT"
  ) +
  scale_x_continuous(breaks = 0:5) +
  labs(
    title    = "Post-treatment ATT by SDG Goal (CS DiD)",
    subtitle = "* = 95% CI excludes zero  |  Red = more missingness  |  Blue = less missingness",
    x        = "Years since autocratization onset",
    y        = "SDG Goal"
  ) +
  theme_minimal(base_size = 12)
```

---

# 8. Robustness: Not-Yet-Treated Controls

```{r cs-overall-nyt, cache=TRUE}
cs_overall_nyt <- att_gt(
  yname         = "prop_sdg_missing",
  tname         = "tname",
  idname        = "idname",
  gname         = "gname",
  xformla       = ~ 1,
  control_group = "notyettreated",
  anticipation  = 1,
  data          = panel_cs %>%
                    filter(!is.na(prop_sdg_missing)) %>%
                    as.data.frame(),
  est_method    = "dr",
  clustervars   = "idname",
  print_details = FALSE
)
```


```{r nyt-dynamic-plot, fig.width=9, fig.height=5}
gt_nyt <- data.frame(
  group = cs_overall_nyt$group,
  time  = cs_overall_nyt$t,
  att   = cs_overall_nyt$att,
  se    = cs_overall_nyt$se
) %>%
  mutate(event_time = time - group)

dynamic_nyt <- gt_nyt %>%
  group_by(event_time) %>%
  summarise(
    att_e = mean(att, na.rm = TRUE),
    se_e  = sqrt(mean(se^2, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  filter(event_time >= -4, event_time <= 5) %>%
  mutate(
    ci_lo = att_e - 1.96 * se_e,
    ci_hi = att_e + 1.96 * se_e
  )

ggplot(dynamic_nyt, aes(x = event_time, y = att_e)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  geom_vline(xintercept = -0.5, linetype = "dashed", colour = "red", alpha = 0.6) +
  geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), fill = "#d6604d", alpha = 0.2) +
  geom_line(colour = "#b2182b", linewidth = 0.8) +
  geom_point(colour = "#b2182b", size = 2.5) +
  scale_x_continuous(breaks = -4:5) +
  labs(
    title    = "CS Event-Study (Robustness): Not-Yet-Treated Controls",
    subtitle = "Overall SDG Missingness | Doubly-robust DiD | Anticipation = 1",
    x        = "Years since autocratization onset",
    y        = "ATT"
  ) +
  theme_minimal(base_size = 13)
```

---

# 9. Optional: SPI High vs. Low Subgroup

```{r spi-models, cache=TRUE}
cs_spi_hi <- tryCatch(
  att_gt(
    yname         = "prop_sdg_missing",
    tname         = "tname",
    idname        = "idname",
    gname         = "gname",
    xformla       = ~ 1,
    control_group = "nevertreated",
    anticipation  = 1,
    data          = panel_cs %>%
                      filter(spi_hilow == "High SPI", !is.na(prop_sdg_missing)) %>%
                      as.data.frame(),
    est_method    = "dr",
    clustervars   = "idname",
    print_details = FALSE
  ),
  error = function(event_time) { message("High SPI failed: ", event_time$message); NULL }
)

cs_spi_lo <- tryCatch(
  att_gt(
    yname         = "prop_sdg_missing",
    tname         = "tname",
    idname        = "idname",
    gname         = "gname",
    xformla       = ~ 1,
    control_group = "nevertreated",
    anticipation  = 1,
    data          = panel_cs %>%
                      filter(spi_hilow == "Low SPI", !is.na(prop_sdg_missing)) %>%
                      as.data.frame(),
    est_method    = "dr",
    clustervars   = "idname",
    print_details = FALSE
  ),
  error = function(event_time) { message("Low SPI failed: ", event_time$message); NULL }
)
```

```{r spi-plots, fig.width=12, fig.height=5}
build_dynamic <- function(cs_obj, label) {
  data.frame(
    group = cs_obj$group, time = cs_obj$t,
    att = cs_obj$att, se = cs_obj$se
  ) %>%
    mutate(event_time = time - group) %>%
    filter(event_time >= -4, event_time <= 5) %>%
    group_by(event_time) %>%
    summarise(
      att_e = mean(att, na.rm = TRUE),
      se_e  = sqrt(mean(se^2, na.rm = TRUE)),
      .groups = "drop"
    ) %>%
    mutate(
      ci_lo = att_e - 1.96 * se_e,
      ci_hi = att_e + 1.96 * se_e,
      label = label
    )
}

if (!is.null(cs_spi_hi) && !is.null(cs_spi_lo)) {
  spi_combined <- bind_rows(
    build_dynamic(cs_spi_hi, "High SPI"),
    build_dynamic(cs_spi_lo, "Low SPI")
  )

  ggplot(spi_combined, aes(x = event_time, y = att_e, colour = label, fill = label)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_vline(xintercept = -0.5, linetype = "dashed", colour = "red", alpha = 0.5) +
    geom_ribbon(aes(ymin = ci_lo, ymax = ci_hi), alpha = 0.15, colour = NA) +
    geom_line(linewidth = 0.8) +
    geom_point(size = 2) +
    scale_x_continuous(breaks = -4:5) +
    facet_wrap(~ label, ncol = 2) +
    labs(
      title = "CS Event-Study by SPI Level: Overall SDG Missingness",
      x     = "Years since autocratization onset",
      y     = "ATT"
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "none")
}
```

---

# 10. Export Results

```{r export}
# Overall dynamic ATT
write_csv(dynamic_overall, "data/output/cs_overall_dynamic_atts.csv")

# Not-yet-treated robustness
write_csv(dynamic_nyt, "data/output/cs_nyt_dynamic_atts.csv")

# Goal-level dynamic ATTs
write_csv(goal_dynamic_df, "data/output/cs_goal_dynamic_atts.csv")

# Group-level ATTs
write_csv(group_overall, "data/output/cs_group_atts.csv")

message("All results exported.")
```

---

# Session Info

```{r session}
sessionInfo()
```
