---
title: "missingness_measure_tests"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# `here` finds the project root (the folder with .git), so paths work whether
# this Rmd is knit or run chunk-by-chunk.
if (!requireNamespace("here", quietly = TRUE)) install.packages("here")
library(here)
library(readr)
library(dplyr)
```

## 1. Load the two inputs

`goal_lvl_dv_data` is the long outcome panel (one row per country x year
x goal). `MAIN_panel_data` is the country x year treatment/control
panel. We read MAIN only to extract from it; we never modify it.

```{r load}
# Outcome panel (long: country x year x goal).
goal_lvl_dv_data <- read_rds(here("data/clean/goal_lvl_dv_data.rds"))

# goal_lvl_dv_data <- read_csv(here("data/clean/goal_lvl_dv_data.csv"),
#                              show_col_types = FALSE)

# Treatment / control panel (country x year). Read-only.
MAIN_panel_data <- read_csv(here("data/output/MAIN_panel_data.csv"),
                            show_col_types = FALSE)

cat("goal_lvl_dv_data:", nrow(goal_lvl_dv_data), "rows |",
    n_distinct(goal_lvl_dv_data$iso3), "countries x",
    n_distinct(goal_lvl_dv_data$year), "years x",
    n_distinct(goal_lvl_dv_data$goal), "goals\n")

cat("MAIN_panel_data:  ", nrow(MAIN_panel_data), "rows |",
    n_distinct(MAIN_panel_data$country_code), "countries x",
    n_distinct(MAIN_panel_data$year), "years\n")
```

## 2. Extract predictors -\> `predictor_data`

Pull only the columns used as treatment, controls, and confounders — NOT
the SDG achievement scores (`sdg1_* … sdg17_*`) or SDR's own per-goal
missingness (`n_miss_SDG*`, `prop_miss_SDG*`); those are alternative
outcomes, not predictors.

Edit the vectors below to add or drop predictors — the extraction
follows.

```{r predictor-data}
# --- merge keys (country_code is ISO-3166 alpha-3, = iso3 in the outcome panel) ---
keys <- c("country_code", "year")

# --- treatment: democracy / regime / autocratization episodes (V-Dem, ERT) ---
treatment_vars <- c(
  "elect_dem", "elect_dem_ert", "lib_dem", "part_dem", "egal_dem", "delib_dem",
  "v2x_liberal",
  "regime_type_2", "regime_type_4", "regime_type_10",
  "regch_event", "regch_genuine",
  "dem_ep", "dem_ep_pre_yr", "dem_ep_start_yr", "dem_ep_end_yr", "dem_ep_id",
  "aut_ep", "aut_ep_pre_yr", "aut_ep_start_yr", "aut_ep_end_yr", "aut_ep_id",
  "G_aut", "D_aut", "aut_cohort"
)

# --- controls / confounders (capacity, wealth, size, violence) ---
control_vars <- c(
  "income_level", "gdp_pc", "log_gdppc", "total_pop", "log_pop",
  "electric_access", "rural_pop_pct", "rd_expenditure_pct",
  "spi_overall",                  # statistical-capacity index — key confounder
  "luminosity", "log_luminosity", # nighttime lights (economic proxy)
  "tax_rev_gdp",
  "poly_vio_events", "any_political_violence", 
  
  # ROBUSTNESS CHECK 
  "prop_sdg_missing" 
)

predictor_cols <- c(keys, treatment_vars, control_vars)

# Fail loudly if a named column is not in MAIN (e.g. a typo or a renamed field).
missing_cols <- setdiff(predictor_cols, names(MAIN_panel_data))
if (length(missing_cols) > 0)
  stop("Not in MAIN_panel_data: ", paste(missing_cols, collapse = ", "))

predictor_data <- MAIN_panel_data %>% select(all_of(predictor_cols))

# Guard: country_code x year must be unique, or the merge would fan out rows.
stopifnot(nrow(predictor_data) ==
          nrow(distinct(predictor_data, country_code, year)))

cat("predictor_data:", nrow(predictor_data), "rows (country x year) x",
    ncol(predictor_data), "cols\n")
```

## 3. Merge onto the long outcome panel -\> `analysis_data`

The predictors are country-year, so they have no goal dimension. Merging
them onto the long outcome panel (`iso3` + `year`) **broadcasts** each
predictor across that country-year's 17 goals — that broadcast IS the
long analysis frame. No pivot of `predictor_data` is needed.

`left_join` keeps every outcome row; countries absent from MAIN get NA
predictors (they drop out of any model on their own). We report the
match rather than silently dropping.

```{r merge}
analysis_data <- goal_lvl_dv_data %>%
  left_join(predictor_data, by = c("iso3" = "country_code", "year" = "year"))

# Match report: which outcome countries found treatment/control data?
matched   <- analysis_data %>% filter(!is.na(elect_dem_ert)) %>% distinct(iso3, geo_area_name)
unmatched <- analysis_data %>% filter(is.na(elect_dem_ert)) %>% distinct(iso3, geo_area_name)

# for compatibility with the original MAIN_panel_data, which uses country_code and country_name.x
analysis_data$country_code <- analysis_data$iso3  
analysis_data$country_name.x <- analysis_data$geo_area_name  

# reorder columns: outcome first, then predictors
analysis_data <- analysis_data %>%
  select(iso3, country_code, geo_area_name, country_name.x, year, scope, goal, starts_with("dv_"), everything())

cat("analysis_data:", nrow(analysis_data), "rows |",
    n_distinct(analysis_data$iso3), "countries\n")
cat("  countries WITH predictors :", nrow(matched), "\n")
cat("  countries WITHOUT (NA)    :", nrow(unmatched),
    " -> rows:", sum(is.na(analysis_data$elect_dem_ert)), "\n")
cat("  complete-case rows (modelable):", sum(!is.na(analysis_data$elect_dem_ert)), "\n")

# Guard: the merge must not have duplicated any outcome rows.
stopifnot(nrow(analysis_data) == nrow(goal_lvl_dv_data))
```

The unmatched countries (not in MAIN_panel_data — mostly micro-states
without a V-Dem/ERT score, plus the four retained non-members):

```{r unmatched}
print(as.data.frame(unmatched[order(unmatched$geo_area_name), ]), row.names = FALSE)
```

## 4. Directional decomposition of democracy

Split year-over-year change in each democracy index into its **decline** and
**improve** components (magnitudes, so backsliding and gains enter as separate
positive regressors). Three indices: `lib_dem` → `ldi`, `elect_dem` → `edi`,
`v2x_liberal` → `liberal`. The `.03` variants keep only changes of at least
0.03 (a meaningful-magnitude threshold).

Computed once per country-year on the distinct set (democracy values are constant
across a country-year's goals), then joined back so every goal row carries them.

```{r dem-decomp}
dem_decomp <- analysis_data %>%
  distinct(iso3, year, lib_dem, elect_dem, v2x_liberal) %>%
  arrange(iso3, year) %>%
  group_by(iso3) %>%
  mutate(
    year_gap = year - lag(year),
    # Liberal Democracy Index (lib_dem -> ldi)
    delta_ldi          = if_else(year_gap == 1, lib_dem - lag(lib_dem), NA_real_),
    decline_ldi        = if_else(delta_ldi < 0,     abs(delta_ldi), 0),
    decline_ldi.03     = if_else(delta_ldi <= -0.03, abs(delta_ldi), 0),
    improve_ldi        = if_else(delta_ldi > 0,      delta_ldi, 0),
    improve_ldi.03     = if_else(delta_ldi >= 0.03,  abs(delta_ldi), 0),
    # Electoral Democracy Index (elect_dem -> edi)
    delta_edi          = if_else(year_gap == 1, elect_dem - lag(elect_dem), NA_real_),
    decline_edi        = if_else(delta_edi < 0,     abs(delta_edi), 0),
    decline_edi.03     = if_else(delta_edi <= -0.03, abs(delta_edi), 0),
    improve_edi        = if_else(delta_edi > 0,      delta_edi, 0),
    improve_edi.03     = if_else(delta_edi >= 0.03,  abs(delta_edi), 0),
    # Liberal component index (v2x_liberal -> liberal)
    delta_liberal      = if_else(year_gap == 1, v2x_liberal - lag(v2x_liberal), NA_real_),
    decline_liberal    = if_else(delta_liberal < 0,     abs(delta_liberal), 0),
    decline_liberal.03 = if_else(delta_liberal <= -0.03, abs(delta_liberal), 0),
    improve_liberal    = if_else(delta_liberal > 0,      delta_liberal, 0),
    improve_liberal.03 = if_else(delta_liberal >= 0.03,  abs(delta_liberal), 0)
  ) %>%
  ungroup()

# Join the decomposition back onto every goal row (by country-year only).
analysis_data <- analysis_data %>%
  left_join(dem_decomp %>% select(iso3, year, year_gap,
                                  starts_with("delta_"), starts_with("decline_"),
                                  starts_with("improve_")),
            by = c("iso3", "year"))
```

## 5. Save

```{r save}
write_csv(analysis_data,  here("data/clean/analysis_data.csv"))
saveRDS(analysis_data,    here("data/clean/analysis_data.rds"))
```

`analysis_data` is the long modelling frame: goal-varying `dv_*`
outcomes plus country-year treatment and controls, ready for `fixest`
with country / goal / year fixed effects and a treatment × goal
interaction.

## 6. Example models

A few core specifications. Primary treatment is the **directional decomposition**
(Section 4): `decline_ldi` (backsliding magnitude) and `improve_ldi` (gains),
each entered contemporaneously and lagged. All use country / goal / year fixed
effects and SEs clustered by country.

**Primary DV: `dv_n_observations_pct_goalbase`** — reporting relative to the
cross-country norm for the goal. `asinh()` handles its right-skew and 0s (log-like,
defined at 0). Modeled with `feols`. `goalbase` is preferred over `pct_baseline`
because it is defined for every country-goal (no dropped late-starters).

```{r models-primary}
library(fixest)

goal <- analysis_data %>% filter(scope == "goal")     # 17-goal panel
ovr  <- analysis_data %>% filter(scope == "overall")  # country-year totals

# 1. Primary: does reporting fall (vs the goal norm) as democracy declines?
#    panel.id lets l() lag within each country-goal series over years.
m_primary <- feols(asinh(dv_n_observations_pct_goalbase)
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  + electric_access
  + I(electric_access^2)
  + rural_pop_pct
  + log_gdppc
  + spi_overall     # SPI mediation (statistical-capacity / ability variable)
  + log_luminosity  # luminosity mediation (ability variable)
  | iso3 + goal + year,
  data = goal, panel.id = ~ iso3^goal + year, 
  cluster = ~iso3)

# 2. Which goals are most sensitive to backsliding? (decline x goal -> per-goal coef)
m_bygoal  <- feols(asinh(dv_n_observations_pct_goalbase) 
                   ~ i(goal, decline_ldi, ref = 1) 
                   | iso3 + goal + year, 
                   data = goal, 
                   cluster = ~iso3)

# 3. Overall country-year reporting (no goal dimension)
m_overall <- feols(asinh(dv_n_observations_pct_goalbase)
  ~ l(decline_ldi, 0:1) 
  + l(improve_ldi, 0:1)
  + electric_access
  + I(electric_access^2)
  + rural_pop_pct
  + log_gdppc
  + spi_overall
  + log_luminosity
  | iso3 + year, data = ovr, 
  panel.id = ~ iso3 + year, 
  cluster = ~iso3)

etable(m_primary, m_overall)
iplot(m_bygoal)  # per-goal sensitivity to democratic decline
```

Robustness: (a) the raw count via Poisson — confirms the result isn't an artifact
of the ratio transform; (b) the country-produced sharp test — if agencies backfill
a country's dropped reporting, the total holds but country-produced data falls.

```{r models-robustness}
# a. same question, native count family
m_count   <- fepois(n_observations
  ~ l(decline_ldi, 0:1) 
  + l(improve_ldi, 0:1)
  + log_gdppc 
  + spi_overall |
  iso3 + goal + year, data = goal, 
  panel.id = ~ iso3^goal + year, 
  cluster = ~iso3)

# b. sharp provenance test (country-produced only; exists as a count)
m_country <- fepois(n_observations_country
  ~ l(decline_ldi, 0:1) 
  + l(improve_ldi, 0:1)
  + log_gdppc 
  + spi_overall |
  iso3 + goal + year, data = goal, 
  panel.id = ~ iso3^goal + year, 
  cluster = ~iso3)

etable(m_count, m_country)
```

> `spi_overall` (statistical capacity) is the key confounder — check its NA
> coverage, as it may shrink the sample.
