---
title: "panel_model_analysis"
output: html_document
---

## Loading prepared panel data & packages
```{r}
source("code/packages.R")

# Data prep 
source("causal_inference_mods/causal_inf_data_prep.R")

# Load data (adjust path if needed)
panel <- readRDS("data/output/MAIN_panel_data.rds")
```

**Source for Luminosity data**: Beyer, R.C., Hu, Y. and Yao, J., 2026. The Bright Side of Heteroskedasticity: Measuring Quarterly Economic Growth from Outer Space.
(link)[https://sites.google.com/site/jiaxiongyao16/nighttime-lights-data?authuser=0] 


## Robustness Check: Assessing Distribution of Key Variables
```{r}
# Distribution of key variables
key_vars <- panel %>%
  select(spi_overall, gdp_pc, total_pop, electric_access, luminosity, rural_pop_pct) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

ggplot(key_vars, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Key Variables", x = "Value", y = "Frequency")
```

Given the largely skewed distribution of population, the log of population is used in all models. We also use the log of luminosity and log of GDP per capita (variable applied as necessary) to reduce skewness in these variables.

## Robustness Check: [Corrected] distribution of vars post-log transformation & imputation
```{r}
# checking distribution of key variables using log-transformed and imputed values
key_vars <- panel %>%
  select(spi_overall, log_gdppc, log_pop, electric_access, log_luminosity, rural_pop_pct) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

ggplot(key_vars, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Key Variables", x = "Value", y = "Frequency")
```


## Addressing NA values in key covariates
```{r}
# # imputed missing values for log_pop with previous year values (carry forward)

# count of missing values in key variables
missing_summary <- panel %>%
  summarise(
    prop_sdg_missing = sum(is.na(prop_sdg_missing)),
    aut_ep = sum(is.na(aut_ep)),
    lib_dem = sum(is.na(lib_dem)),
    elect_dem = sum(is.na(elect_dem)),
    v2x_liberal = sum(is.na(v2x_liberal)),
    log_gdppc = sum(is.na(log_gdppc)),
    log_pop = sum(is.na(log_pop)),
    electric_access = sum(is.na(electric_access)),
    log_luminosity = sum(is.na(log_luminosity)),
    rural_pop_pct = sum(is.na(rural_pop_pct)),
    spi_overall = sum(is.na(spi_overall)),
    tax_rev_gdp = sum(is.na(tax_rev_gdp))
  )
print(missing_summary)

# Table of countries with remaining NA values in key variables (after log_pop imputation)
na_values_by_country <- panel %>%
  group_by(country_code, country_name.x) %>%
  summarise(across(c(log_gdppc, log_pop, log_luminosity, electric_access, spi_overall, tax_rev_gdp, rural_pop_pct),
                    ~ sum(is.na(.x))), .groups = "drop") %>%
  filter(if_any(c(log_gdppc, log_pop, log_luminosity, electric_access, spi_overall, tax_rev_gdp, rural_pop_pct),
                ~ .x > 0)) %>%
  janitor::adorn_totals("row")

# Export the actual data [TO APPENDIX]
write_csv(na_values_by_country, "results_csv/na_values_by_country.csv")

# Table of country NA values per variable 
var_labels_na <- c(
  "Log GDP per Capita" = "log_gdppc",
  "Log Population" = "log_pop",
  "Log Nighttime Luminosity" = "log_luminosity",
  "Electricity Access (%)" = "electric_access",
  "Statistical Performance Index" = "spi_overall",
  "Tax Revenue (% of GDP)" = "tax_rev_gdp", 
   "Rural Population (%)" = "rural_pop_pct"
)

na_table <- na_values_by_country %>%
  mutate(`#` = if_else(country_code == "Total", NA_integer_, row_number())) %>%
  relocate(`#`) %>%
  rename(Country = country_name.x, `Iso3 Code` = country_code) %>%
  rename(!!!var_labels_na)

set_flextable_defaults(na_str = "NA")
highlight_row <- which(na_table$Country %in% c("North Korea", "Eritrea", "Venezuela", "Cuba", "South Sudan"))
na_ft <- na_table %>% flextable()

na_ft <- na_ft %>%
  theme_apa() %>%
  padding(padding.top = 1, padding.bottom = 1, part = "body") %>%
  bg(i = highlight_row, bg = "#FFF3CD", part = "body") %>%
  add_header_lines(values = "Countries with remaining missing values by variable (2015-2023)") %>%
  bold(part = "header", i = 1) %>%
  bold(part = "body", i = nrow(na_table)) %>%
  add_footer_lines(
    "Note. Cell values indicate the number of country-year observations with missing data for each control variable considered in this study, within the 2015-2023 window. Values reflect counts after log population imputation. The final row reports totals across all countries. This table focuses on control variables, as all democracy predictors contain zero missing values across the full sample."
  ) %>%
  italic(part = "footer") %>%
  fontsize(size = 10, part = "footer") %>%
  font(fontname = "Times New Roman", part = "footer") %>%
  autofit()

save_as_image(na_ft, path = "figures/causal_inference/na_values_by_country_table.png")

```

**total_pop** has the most missing values, specifically across all countries in 2023, which is likely due to the fact that the World Bank's population data for 2023 may not have been fully updated or released at the time of data collection. 

Strategy: Population data for 2023 was unavailable from the World Bank at the time of data collection; 2022 values were carried forward for this variable. Given the low year-over-year volatility of (log) population, this is unlikely to meaningfully affect estimates. 


## Robustness check: VIF multicollinearity across model variables 
```{r}
# VIF 1 (All)
vif_model <- lm(prop_sdg_missing ~ lib_dem + log_gdppc + log_pop + spi_overall + electric_access + log_luminosity + rural_pop_pct,
                 data = panel %>% select(prop_sdg_missing, lib_dem, log_gdppc, log_pop, spi_overall, electric_access, log_luminosity, rural_pop_pct))

# VIF 2 (without log_gdppc)
vif2_model <- lm(prop_sdg_missing ~ lib_dem + log_pop + spi_overall + electric_access + log_luminosity + rural_pop_pct,
                  data = panel %>% select(prop_sdg_missing, lib_dem, log_pop, spi_overall, electric_access, log_luminosity, rural_pop_pct))

# # VIF 3 (without log_pop)
vif3_model <- lm(prop_sdg_missing ~ lib_dem + log_gdppc + spi_overall + electric_access + log_luminosity + rural_pop_pct,
                  data = panel %>% select(prop_sdg_missing, lib_dem, log_gdppc, spi_overall, electric_access, log_luminosity, rural_pop_pct))

# # VIF 4 (without SPI)
vif4_model <- lm(prop_sdg_missing ~ lib_dem + log_gdppc + log_pop + electric_access + log_luminosity + rural_pop_pct,
                  data = panel %>% select(prop_sdg_missing, lib_dem, log_gdppc, log_pop, electric_access, log_luminosity, rural_pop_pct)) 

# Create a table of VIF results for all four models
var_labels <- c(
  lib_dem = "Liberal Democracy Index",
  elect_dem = "Electoral Democracy Index",
  spi_overall = "Statistical Performance Index",
  log_pop = "Log Population",
  log_gdppc = "Log GDP per Capita",
  log_luminosity = "Log Nighttime Luminosity",
  electric_access = "Electricity Access (%)",
  rural_pop_pct = "Rural Population (%)"
)

vif_table <- reduce(
  list(
    vif(vif_model)  %>% enframe(name = "Variable", value = "VIF 1: All "),
    vif(vif2_model) %>% enframe(name = "Variable", value = "VIF 2: No GDP "),
    vif(vif3_model) %>% enframe(name = "Variable", value = "VIF 3: No Pop "),
    vif(vif4_model) %>% enframe(name = "Variable", value = "VIF 4: No SPI ")
  ),
  full_join, by = "Variable"
) %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))

n_row <- tibble(
  Variable = "N (Observations)",
  `VIF 1: All ` = nobs(vif_model),
  `VIF 2: No GDP ` = nobs(vif2_model),
  `VIF 3: No Pop ` = nobs(vif3_model),
  `VIF 4: No SPI ` = nobs(vif4_model)
)

vif_table <- bind_rows(vif_table, n_row)

set_flextable_defaults(na_str = "NA")

vif_ft <- vif_table %>%
  flextable()

vif_ft <- vif_ft %>%
  theme_apa() %>%
  add_header_lines(values = "VIF & country-year count comparison across four model specifications") %>%
  padding(padding.top = 1, padding.bottom = 1, part = "body") %>%
  bold(part = "header", i = 1) %>%
  bold(part = "body", i = nrow(vif_table)) %>%
  colformat_double(j = c("VIF 1: All ", "VIF 2: No GDP ", "VIF 3: No Pop ", "VIF 4: No SPI "), digits = 2) %>%
  colformat_double(i = ~ Variable == "N (Observations)",
                    j = c("VIF 1: All ", "VIF 2: No GDP ", "VIF 3: No Pop ", "VIF 4: No SPI "),
                    digits = 0, big.mark = ",") %>%
  add_footer_lines("Note. Variance inflation factors (VIFs) were calculated from regressions of overall SDG missingness on all covariates included in each specification. Columns 2 through 4 sequentially remove one covariate at a time (population, GDP, and SPI, respectively) to assess whether any single covariate is driving inflated VIFs among the others. VIF scores < 5 signify low collinearity; scores \u2265 5 signify moderate collinearity; and scores \u2265 10 exhibit severe collinearity.") %>%
  italic(part = "footer") %>%
  fontsize(size = 10, part = "footer") %>%
  font(fontname = "Times New Roman", part = "footer") %>%
  autofit()

save_as_image(vif_ft, path = "figures/causal_inference/vif1-4_table_twfe.png")
```

## Robustness Check: Correlation matrix of control variables
```{r}
# matrix of variables and correlations
cor_data <- panel %>% select(prop_sdg_missing, lib_dem, elect_dem, v2x_liberal, 
                             log_gdppc, log_pop, spi_overall, electric_access, 
                             log_luminosity, rural_pop_pct, tax_rev_gdp) %>% na.omit()
cor_matrix <- cor(cor_data)
cor_matrix %>% round(2) %>% knitr::kable(
  caption = "Correlation matrix for control variables")

# # All covariates used in models
# source("descriptive_stats/corr_matrix_key_vars.R")
```

*NOTES ABOUT VARIABLES:*
- Took out R&D expenditure because it has a lot of NAs and is highly correlated with GDP per capita.
- SPI is missing all of 2015, but is kept in the dataset for now because it is a key control variable. Models are run with and without it to check robustness.

## Directional Decomposition: LDI, EDI & LCI 
```{r}
panel$year <- as.numeric(as.character(panel$year))
panel$country_code <- as.character(panel$country_code)

panel <- panel %>%
  arrange(country_code, year) %>%
  group_by(country_code) %>%
  mutate(
    year_gap = year - dplyr::lag(year),
    delta_ldi = ifelse(year_gap == 1, lib_dem - dplyr::lag(lib_dem), NA_real_),
    decline_ldi = ifelse(delta_ldi < 0, abs(delta_ldi), 0),
    decline_ldi.03 = ifelse(delta_ldi <= -0.03, abs(delta_ldi), 0),
    improve_ldi = ifelse(delta_ldi > 0, delta_ldi, 0),
    improve_ldi.03 = ifelse(delta_ldi >= 0.03, abs(delta_ldi), 0),
    delta_edi = ifelse(year_gap == 1, elect_dem - dplyr::lag(elect_dem), NA_real_),
    decline_edi = ifelse(delta_edi < 0, abs(delta_edi), 0),
    decline_edi.03 = ifelse(delta_edi <= -0.03, abs(delta_edi), 0),
    improve_edi = ifelse(delta_edi > 0, delta_edi,  0),
    improve_edi.03 = ifelse(delta_edi >= 0.03, abs(delta_edi), 0),
    delta_liberal = ifelse(year_gap == 1, v2x_liberal - dplyr::lag(v2x_liberal), NA_real_),
    decline_liberal = ifelse(delta_liberal < 0, abs(delta_liberal), 0),
    decline_liberal.03 = ifelse(delta_liberal <= -0.03, abs(delta_liberal), 0),
    improve_liberal = ifelse(delta_liberal > 0, delta_liberal, 0),
    improve_liberal.03 = ifelse(delta_liberal >= 0.03, abs(delta_liberal), 0),
    # Creating varibles detecting high-magnitude changes (>= 0.3) in sdg missingness
  ) %>%
  ungroup()
```

## Some Visual Analysis of Backsliding and Democratizing Magnitudes
```{r}
## Histograms of backsliding and democratizing magnitudes (liberal democracy index)

# Backsliding magnitudes (country-years with decline_ldi > 0)
backsliding_ldi_vals <- panel$decline_ldi[panel$decline_ldi != 0]

hist(
  backsliding_ldi_vals,
  main = "Histogram of Backsliding Magnitudes (lib_dem)",
  xlab = "Year-to-year decline in lib_dem",
  ylab = "Number of country-years",
  col  = "lightgray",
  border = "black"
)

# Democratizing magnitudes (country-years with improve_ldi > 0)
democratizing_ldi_vals <- panel$improve_ldi[panel$improve_ldi != 0]

hist(
  democratizing_ldi_vals,
  main = "Histogram of Democratizing Magnitudes (lib_dem)",
  xlab = "Year-to-year improvement in lib_dem",
  ylab = "Number of country-years",
  col  = "lightgray",
  border = "black"
)

## Three-by-two table of country-year counts by type and magnitude 
# [Directional distribution and magnitude of LDI changes across 1,3XX country-years (2015–2023)]

# Backsliding: any decline vs decline >= 0.3
backsliding_any   <- sum(panel$decline_ldi   != 0, na.rm = TRUE)
backsliding_0.03   <- sum(panel$decline_ldi.03 != 0, na.rm = TRUE)

# Democratizing: any improvement vs improvement >= 0.3
democratizing_any <- sum(panel$improve_ldi   != 0, na.rm = TRUE)
democratizing_0.03 <- sum(panel$improve_ldi.03 != 0, na.rm = TRUE)

# Stable: no decline and no improvement (any magnitude),
# and no decline >= 0.3 and no improvement >= 0.3
stable_any <- sum(
  panel$decline_ldi   == 0 &
  panel$improve_ldi   == 0,
  na.rm = TRUE
)

stable_0.03 <- sum(
  panel$decline_ldi.03 == 0 &
  panel$improve_ldi.03 == 0,
  na.rm = TRUE
)

# Compose 3x2 table
ldi_mag_table <- rbind(
  backsliding = c(any_change = backsliding_any, change_ge_0.03 = backsliding_0.03),
  democratizing = c(any_change = democratizing_any, change_ge_0.03 = democratizing_0.03),
  stable = c(any_change = stable_any, change_ge_0.03 = stable_0.03))

ldi_mag_table
```

## PART 1: Two‑way fixed effects (TWFE) [Baseline Models] [LDI Decomposed]
```{r}
# ---------------------- lib_dem -> overall SDG missingness ---------------------

# M1: Reduced Dynamic TWFE (just to see raw relationship)
m1_twfe_a <- feols(
  prop_sdg_missing 
  ~ decline_ldi + improve_ldi 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m1_twfe_a)

# M2: Reduced Dynamic TWFE + Lags
m2_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m2_twfe_a)

# M3: Full Dynamic TWFE + Lags + Controls (SPI included as a control)
m3_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(decline_ldi, 0:1)
  #+ spi_overall
  #+ log_luminosity
  #+ log_pop
  #+ log_gdppc
  + electric_access 
  + I(electric_access^2)
  + rural_pop_pct 
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m3_twfe_a)


# M4: Dynamic TWFE + Lags + Confounders + economic preconditions/consequences (GDP robustness)
m4_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  + electric_access 
  + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc # Adding economic conditions 
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m4_twfe_a)

# M5: Dynamic TWFE + Lags + Confounders + economic preconditions/consequences (GDP robustness)
m5_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  + electric_access 
  + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc
  + spi_overall # adding SPI mediation (ability variable)
  + log_luminosity # luminosity mediation (ability variable)
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m5_twfe_a)

# ==============================================================================

# M5a/b: Full Dynamic TWFE + Lags + Controls + Heterogeneity
# (Ability vs willingness: aut_ep x income_level, aut_ep x regime_type)

# Checking for heterogeneity in income_level and regime_type_4 over time (i.e., whether any countries change income_level or regime_type_4 over time)
table(panel$country_code, panel$income_level)
panel %>% distinct(country_code, income_level) %>% count(country_code) %>% filter(n > 1) %>% nrow()

table(panel$country_code, panel$regime_type_4)
panel %>% distinct(country_code, regime_type_4) %>% count(country_code) %>% filter(n > 1) %>% nrow()

# M5a: Heterogeneity by income
m5a_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  + decline_ldi * income_level
  + electric_access + I(electric_access^2)
  + rural_pop_pct
  + log_gdppc
  + spi_overall
  + log_luminosity
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m5a_twfe_a)

# M5b: Heterogeneity by regime type
m5b_twfe_a <- feols(
  prop_sdg_missing 
  ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)
  + decline_ldi * regime_type_4
  + log_gdppc
  + electric_access + I(electric_access^2)
  + rural_pop_pct
  + spi_overall
  + log_luminosity
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m5b_twfe_a)
# ==============================================================================

models_ldi <- list(
  "M1 Reduced" = m1_twfe_a,
  "M2 + Lags" = m2_twfe_a,
  "M3 + Structural Confounders" = m3_twfe_a,
  "M4 + Economic Conditions" = m4_twfe_a,
  "M5 + Ability Mediators" = m5_twfe_a,
  "M5a \u00d7 Income" = m5a_twfe_a,
  "M5b \u00d7 Regime" = m5b_twfe_a)

extra_rows <- data.frame(
  term = c("N (Years)", "N (Countries)"),

  "M1 Reduced"                  = as.character(c(m1_twfe_a$fixef_sizes["year"],
                                                  m1_twfe_a$fixef_sizes["country_code"])),
  "M2 + Lags"                   = as.character(c(m2_twfe_a$fixef_sizes["year"],
                                                  m2_twfe_a$fixef_sizes["country_code"])),
  "M3 + Structural Confounders" = as.character(c(m3_twfe_a$fixef_sizes["year"],
                                                  m3_twfe_a$fixef_sizes["country_code"])),
  "M4 + Economic Conditions"    = as.character(c(m4_twfe_a$fixef_sizes["year"],
                                                  m4_twfe_a$fixef_sizes["country_code"])),
  "M5 + Ability Mediators"      = as.character(c(m5_twfe_a$fixef_sizes["year"],
                                                  m5_twfe_a$fixef_sizes["country_code"])),
  "M5a × Income"                = as.character(c(m5a_twfe_a$fixef_sizes["year"],
                                                  m5a_twfe_a$fixef_sizes["country_code"])),
  "M5b × Regime"                = as.character(c(m5b_twfe_a$fixef_sizes["year"],
                                                  m5b_twfe_a$fixef_sizes["country_code"])),

  check.names = FALSE,
  stringsAsFactors = FALSE
)
attr(extra_rows, "position") <- "gof_start"

ldi_gt <- modelsummary(
  models_ldi,
  coef_map = c(
    "decline_ldi"                       = "Decline in LDI",
    "l(decline_ldi, 0)"                  = "Decline in LDI",
    "l(decline_ldi, 1)"                  = "Decline in LDI (t-1)",
    "improve_ldi"                        = "Improvement in LDI",
    "l(improve_ldi, 0)"                  = "Improvement in LDI",
    "l(improve_ldi, 1)"                  = "Improvement in LDI (t-1)",
    "electric_access"                    = "Electricity Access (%)",
    "I(I(electric_access^2))"            = "Electricity Access (%)\u00b2",
    "rural_pop_pct"                      = "Rural Population (%)",
    "log_gdppc"                          = "Log GDP per Capita",
    "spi_overall"                        = "Statistical Performance Index",
    "log_luminosity"                     = "Log Nighttime Luminosity",
    "income_levelLM"                     = "Lower-Middle Income",
    "income_levelUM"                     = "Upper-Middle Income",
    "income_levelH"                      = "High Income",
    "decline_ldi:income_levelLM"         = "Decline in LDI \u00d7 Lower-Middle Income",
    "decline_ldi:income_levelUM"         = "Decline in LDI \u00d7 Upper-Middle Income",
    "decline_ldi:income_levelH"          = "Decline in LDI \u00d7 High Income",
    "regime_type_41"                     = "Regime Type 1",
    "regime_type_42"                     = "Regime Type 2",
    "regime_type_43"                     = "Regime Type 3",
    "decline_ldi:regime_type_41"         = "Decline in LDI \u00d7 Regime Type 1",
    "decline_ldi:regime_type_42"         = "Decline in LDI \u00d7 Regime Type 2",
    "decline_ldi:regime_type_43"         = "Decline in LDI \u00d7 Regime Type 3"
  ),
  gof_map = list(
    list(raw = "nobs", clean = "N (Observations)", fmt = 0),
    list(raw = "ngroups", clean = "N (Groups)", fmt = 0),
    list(raw = "r2.within", clean = "Within R\u00b2", fmt = 3),
    list(raw = "r2.within.adjusted", clean = "Adjusted Within R\u00b2", fmt = 3),
    list(raw = "rmse", clean = "RMSE", fmt = 3)
  ),
  stars = TRUE,
  title = "TWFE Models: Liberal Democracy Index (Decomposed) and SDG Data Missingness",
  output = "gt",
  add_rows = extra_rows
)

ldi_gt <- ldi_gt %>%
tab_options(
  data_row.padding = px(2),
  table.border.top.color = "black",
  table.border.top.width = px(2),
  table.border.bottom.color = "black",
  table.border.bottom.width = px(0),
  heading.border.bottom.color = "black",
  heading.border.bottom.width = px(1),
  column_labels.border.top.color = "black",
  column_labels.border.top.width = px(1),
  column_labels.border.bottom.color = "black",
  column_labels.border.bottom.width = px(2),
  table_body.border.bottom.color = "black",
  table_body.border.bottom.width = px(2),
  footnotes.border.bottom.color = "white",
  footnotes.border.bottom.width = px(0),
  table_body.hlines.color = "white",
  table.font.names = "Times New Roman",
  table.font.size = px(15)
  ) %>%
  tab_style(
    style = cell_text(weight = "bold", align = "center"),
    locations = cells_title()
  ) %>%
  tab_style(
    style = cell_text(style = "italic", size = px(12)),
    locations = cells_footnotes()
  ) %>%
  cols_width(
    " " ~ px(210),
    "M1 Reduced" ~ px(110),
    "M2 + Lags" ~ px(110),
    "M3 + Structural Confounders" ~ px(110),
    "M4 + Economic Conditions" ~ px(110),
    "M5 + Ability Mediators" ~ px(110),
    "M5a \u00d7 Income" ~ px(110),
    "M5b \u00d7 Regime" ~ px(110)
  ) %>% 
  cols_label(
  "M1 Reduced" = md("**M1**<br>Reduced"),
  "M2 + Lags" = md("**M2**<br>+ Lags"),
  "M3 + Structural Confounders" = md("**M3**<br>+ Structural Confounders"),
  "M4 + Economic Conditions" = md("**M4**<br>+ Economic Conditions"),
  "M5 + Ability Mediators" = md("**M5**<br>+ Ability Mediators"),
  "M5a \u00d7 Income" = md("**M5a**<br>\u00d7 Income"),
  "M5b \u00d7 Regime" = md("**M5b**<br>\u00d7 Regime")
  ) %>%
  tab_style(
  style = cell_text(align = "center"),
  locations = cells_column_labels(columns = everything())
  ) %>%
  tab_source_note(source_note = "Note. Cell entries report coefficient estimates with standard errors clustered by country in parentheses below. All models include country and year fixed effects. The dependent variable is the proportion of SDG indicators missing overall.")

gtsave(ldi_gt, "figures/causal_inference/twfe_lib_dem_table.png")
```


# PART 2: Two‑way fixed effects (TWFE) [Goal-Level Models]
Selected Model Specification: M3: + Structural Confounders (electricity access, rural population share)
```{r}
## FITTING MODEL 3 SPECIFICATION

#sub_panel <- subset(panel, income_level %in% c("UM", "H")) # CONSIDERING ONLY HIGH INCOME [HETEROGENEITY]
dv_names <- paste0("prop_miss_SDG", 1:17)

goal_models <- map(dv_names, function(dv) {
  feols(
    as.formula(paste0(
      dv, " ~ l(decline_ldi, 0:1) + l(improve_ldi, 0:1)",
      "+ electric_access + I(electric_access^2) + rural_pop_pct",
      #"+ log_gdppc",
      #"+ spi_overall + log_luminosity",
      " | country_code + year"
    )),
    panel.id = ~ country_code + year,
    data = panel, #sub_panel, 
    cluster = ~ country_code
  )
})
names(goal_models) <- dv_names

## EXTRACTION
goal_results <- map_dfr(names(goal_models), function(dv) {
  m <- goal_models[[dv]]
  co <- tidy(m)
  
  get_est <- function(term) co$estimate[co$term == term]
  get_se  <- function(term) co$std.error[co$term == term]
  get_p   <- function(term) co$p.value[co$term == term]
  
  tibble(
    goal          = dv,
    decline_t     = get_est("l(decline_ldi, 0)"),
    decline_t_se  = get_se("l(decline_ldi, 0)"),
    decline_t_p   = get_p("l(decline_ldi, 0)"),
    decline_l1    = get_est("l(decline_ldi, 1)"),
    decline_l1_se = get_se("l(decline_ldi, 1)"),
    decline_l1_p  = get_p("l(decline_ldi, 1)"),
    improve_t     = get_est("l(improve_ldi, 0)"),
    improve_t_se  = get_se("l(improve_ldi, 0)"),
    improve_t_p   = get_p("l(improve_ldi, 0)"),
    improve_l1    = get_est("l(improve_ldi, 1)"),
    improve_l1_se = get_se("l(improve_ldi, 1)"),
    improve_l1_p  = get_p("l(improve_ldi, 1)"),
    n             = nobs(m),
    r2_within     = r2(m, "wr2")
  )
})

# STARS AND FORMAT
add_stars <- function(est, p) {
  s <- case_when(p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", p < .1 ~ "+", TRUE ~ "")
  paste0(sprintf("%.3f", est), s)
  }

goal_table <- goal_results %>%
  mutate(
    Decline_t = paste0(
      add_stars(decline_t, decline_t_p), " ",
      "(", sprintf("%.3f", decline_t_se), ")"
    ),
    Decline_l1 = paste0(
      add_stars(decline_l1, decline_l1_p), " ",
      "(", sprintf("%.3f", decline_l1_se), ")"
    ),
    Improve_t = paste0(
      add_stars(improve_t, improve_t_p), " ",
      "(", sprintf("%.3f", improve_t_se), ")"
    ),
    Improve_l1 = paste0(
      add_stars(improve_l1, improve_l1_p), " ",
      "(", sprintf("%.3f", improve_l1_se), ")"
    )
  ) %>%
  transmute(
    Goal          = goal,
    `Decline (t)`   = Decline_t,
    `Decline (t-1)` = Decline_l1,
    `Improve (t)`   = Improve_t,
    `Improve (t-1)` = Improve_l1,
    N             = n,
    `Within R²`   = sprintf("%.3f", r2_within)
  )

## APA STYLE TABLE
set_flextable_defaults(na_str = "NA")

goal_ft <- goal_table %>%
  flextable() %>%
  theme_apa() %>%
  padding(padding.top = 1, padding.bottom = 1, part = "body") %>%
  add_header_lines(values = "M3 Specification: Goal-Level Susceptibility to Democratic Decline (SDG1–SDG17)") %>%
  bold(part = "header", i = 1) %>%
  add_footer_lines(
    "Note. All models estimate M3 (dynamic lags plus controls: SPI, log population, electricity access, log luminosity, rural population share). Country and year fixed effects included; SEs clustered by country. + p < .1, * p < .05, ** p < .01, *** p < .001."
  ) %>%
  italic(part = "footer") %>%
  fontsize(size = 10, part = "footer") %>%
  font(fontname = "Times New Roman", part = "footer") %>%
  autofit()

save_as_image(goal_ft, path = "figures/causal_inference/twfe_goal_level_m3_table.png")

# save panel data with goal-level missingness for future use
write_csv(panel, "data/output/MAIN_panel_data.csv")
saveRDS(panel, "data/output/MAIN_panel_data.rds")
```


# [APPENDIX] PART 3: Two‑way fixed effects (TWFE) [Baseline Models] [EDI + LCI Decomposed]
```{r}
# Preliminary VIF Check [Includes: elect_dem and v2x_liberal]

# # VIF 5 
# vif5_model <- lm(prop_sdg_missing ~ 
#                  elect_dem 
#                  + v2x_liberal
#                  + log_gdppc 
#                  #+ log_pop 
#                  + electric_access 
#                  + log_luminosity 
#                  + rural_pop_pct
#                  + spi_overall, 
#                  data = vif_data)
# vif(vif5_model)

# ---------------------- elect_dem -> overall SDG missingness ---------------------

# M1: Reduced Dynamic TWFE (just to see raw relationship)
m1_twfe_b <- feols(
  prop_sdg_missing 
  ~ decline_edi + improve_edi 
  + decline_liberal + improve_liberal
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m1_twfe_b)

# M2: Reduced Dynamic TWFE + Lags
m2_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m2_twfe_b)

# M3: Full Dynamic TWFE + Lags + structural confounders (SPI included as a control)
m3_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m3_twfe_b)

# M4: Full Dynamic TWFE + Lags + Economic Conditions (GDP robustness)
m4_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m4_twfe_b)

# M5: Full Dynamic TWFE + Lags + GDP + Ability Mediators (SPI, Luminosity)
m5_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc
  + spi_overall
  + log_luminosity
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m5_twfe_b)

# M5: Full Dynamic TWFE + Lags + Controls + Heterogeneity
# (Ability vs willingness: aut_ep x income_level, aut_ep x regime_type)

# M5a: Heterogeneity by income
m4a_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  + decline_edi * income_level
  + spi_overall
  + log_gdppc
  + log_luminosity
  + electric_access + I(electric_access^2)
  + rural_pop_pct
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m5a_twfe_b)

# M5b: Heterogeneity by regime type
m5b_twfe_b <- feols(
  prop_sdg_missing 
  ~ l(decline_edi, 0:1) + l(improve_edi, 0:1)
  + l(decline_liberal, 0:1) + l(improve_liberal, 0:1)
  + decline_edi * regime_type_4
  + spi_overall
  + log_gdppc
  + log_luminosity
  + electric_access + I(electric_access^2)
  + rural_pop_pct
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m5b_twfe_b)

```

# [APPENDIX] PART 4: Two‑way fixed effects (TWFE) [Overall - Baseline Models] [aut_ep]
```{r}
# ---------------------- aut_ep -> overall SDG missingness ---------------------

# M1: Reduced Dynamic TWFE (just to see raw relationship)
m1_twfe_c <- feols(
  prop_sdg_missing 
  ~ aut_ep 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m1_twfe_c)

# M2: Reduced Dynamic TWFE + Lags
m2_twfe_c <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:2)
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m2_twfe_c)

# M3: Full Dynamic TWFE + Lags + Controls (SPI included as a control)
m3_twfe_c <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:2)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m3_twfe_c)

# M4
m4_twfe_c <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:2)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc  
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m4_twfe_c)

# M5 
m5_twfe_c <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:2)
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc  
  + spi_overall
  + log_luminosity
  | country_code + year,
  panel.id = ~ country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m5_twfe_c)


# M4: Full Dynamic TWFE + Lags + Controls + Heterogeneity
# (Ability vs willingness: aut_ep x income_level, aut_ep x regime_type)

# Checking for heterogeneity in income_level and regime_type_4 over time (i.e., whether any countries change income_level or regime_type_4 over time)
table(panel$country_code, panel$income_level)
panel %>% distinct(country_code, income_level) %>% count(country_code) %>% filter(n > 1) %>% nrow()

table(panel$country_code, panel$regime_type_4)
panel %>% distinct(country_code, regime_type_4) %>% count(country_code) %>% filter(n > 1) %>% nrow()

# M4a: Heterogeneity by income
m4a_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1)
  + aut_ep * income_level
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc  
  + spi_overall
  + log_luminosity
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m4a_twfe)

# M4b: Heterogeneity by regime type
m4b_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:2)
  + aut_ep * regime_type_4
  + electric_access + I(electric_access^2)
  + rural_pop_pct 
  + log_gdppc  
  + spi_overall
  + log_luminosity
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code
)

summary(m4b_twfe)
```

# ============== ============== ============== ============== ============== ==============
# COMPLEMENTARY MODELS: Heterogeneity by SPI 
```{r}
# (X1) dynamic + SPI heterogeneity (appendix): Ability vs willingness: aut_ep × income_level
mx1_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1) * spi_overall 
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code)

summary(mx1_twfe)

```

