---
title: "panel_model_analysis"
output: html_document
---

## Loading & Merging Data 
```{r}
library(tidyverse)
library(devtools)
library(fixest)   # for TWFE
library(did)      # for Callaway-Sant'Anna DiD
library(car)
library(readr)
library(janitor)
library(WDI) # call WDI package for Vars 

# Load data (adjust path if needed)
# df <- read_csv("sdg_democracy_paneldata.csv")
df <- readRDS("data/output/sdg_democracy_paneldata.rds")

# Quickly Adding merge of SPI data
url <- 
  "https://raw.githubusercontent.com/worldbank/SPI/refs/heads/master/03_output_data/SPI_index.csv"
spi <- read_csv(url) %>% 
  dplyr::select(country, date, SPI.INDEX, iso3c) %>% 
  rename(country_name = country, country_code = iso3c, year = date, spi_overall = SPI.INDEX)

# merging spi into main panel dataset
panel <- df %>% left_join(spi, by = c("country_code", "year"))

# Quickly Adding Luminosity Data (for potential future use) 
luminosity <- read_csv("data/input/csvs_other_indicators/nightlight_1992_2023_complete.csv") %>% 
  filter(year >= 2015) %>% 
  rename(country_code = iso, luminosity = nlsum)

# merging luminosity into main panel dataset
panel <- panel %>% left_join(luminosity, by = c("country_code", "year"))

# Quickly Adding episode id for autocratization and democratization
ert <- read.csv("data/input/ert.csv") %>% 
  filter(year >= 2000) %>% 
  dplyr::select(country_text_id, year, dem_ep_id, aut_ep_id) %>% 
  rename(country_code = country_text_id)

# merging episode id into main panel dataset
panel <- panel %>% left_join(ert, by = c("country_code", "year"))

# Quickly Adding tax revenue data (for potential future use) [WDI]
tax_rev_gdp <- WDI(country = "all", indicator = "GC.TAX.TOTL.GD.ZS", start = 2000, end = NULL)

# merging tax revenue data into main panel dataset
panel <- panel %>% left_join(tax_rev_gdp %>% 
                               rename(country_code = iso3c, tax_rev_gdp = GC.TAX.TOTL.GD.ZS), 
                             by = c("country_code", "year"))
```

**Source for Luminosity data**: Beyer, R.C., Hu, Y. and Yao, J., 2026. The Bright Side of Heteroskedasticity: Measuring Quarterly Economic Growth from Outer Space.
(link)[https://sites.google.com/site/jiaxiongyao16/nighttime-lights-data?authuser=0] 

## Assessing Distribution of Key Variables
```{r}
# Distribution of key variables
key_vars <- panel %>%
  select(spi_overall, log_gdppc, log_pop, electric_access, luminosity, rural_pop_pct) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

ggplot(key_vars, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Key Variables", x = "Value", y = "Frequency")
```



## Data Cleaning and Variable Setup
```{r}
# Basic cleaning / type setting -----------------------------------------------
panel <- panel %>%
  filter(year >= 2015) %>%
  mutate(
    # Treatment cohort: first year of autocratization (0 if never treated)
    aut_cohort = if_else(aut_ep == 1, aut_ep_start_yr, 0L),
    log_luminosity = log(luminosity + 1),
    # Numeric ID required by did::att_gt()
    country_id = as.numeric(as.factor(country_code)))

panel <- panel %>% mutate(across(
    c(country_code, year, regime_type_2, regime_type_4, regime_type_10, 
      aut_ep, dem_ep, regch_event, regch_genuine, aut_cohort), as.factor))

panel$regime_type_4 <- factor(panel$regime_type_4, levels = c("0", "1", "2", "3"))
panel$income_level <- factor(panel$income_level, levels = c("L", "LM", "UM", "H"))
```
**log_pop** has the most missing values, specifically across all countries in 2023, which is likely due to the fact that the World Bank's population data for 2023 may not have been fully updated or released at the time of data collection. 

Strategy: Population data for 2023 was unavailable from the World Bank at the time of data collection; 2022 values were carried forward for this variable. Given the low year-over-year volatility of log population, this is unlikely to meaningfully affect estimates.

## Addressing NA values in key covariates
```{r}
# impute missing values for log_pop with previous year values (carry forward)
panel <- panel %>% 
  group_by(country_code) %>% 
  mutate(
    log_pop = if_else(year == 2023 & is.na(log_pop),
    log_pop[year == 2022], log_pop)
    ) %>% 
  ungroup()

# count of missing values in key variables
missing_summary <- panel %>%
  summarise(
    prop_sdg_missing = sum(is.na(prop_sdg_missing)),
    aut_ep = sum(is.na(aut_ep)),
    lib_dem = sum(is.na(lib_dem)),
    log_gdppc = sum(is.na(log_gdppc)),
    log_pop = sum(is.na(log_pop)),
    electric_access = sum(is.na(electric_access)),
    log_luminosity = sum(is.na(log_luminosity)),
    rural_pop_pct = sum(is.na(rural_pop_pct)),
    spi_overall = sum(is.na(spi_overall)),
    tax_rev_gdp = sum(is.na(tax_rev_gdp))
  )
print(missing_summary)

# Table of countries with remaining NA values in key variables (after log_pop imputation)
panel %>% 
  group_by(country_code, country_name.x) %>% 
  summarise(
    across(c(log_gdppc, log_pop, log_luminosity, electric_access, tax_rev_gdp), ~ sum(is.na(.x))),
    .groups = "drop") %>% 
  filter(if_any(c(log_gdppc, log_pop, log_luminosity, electric_access, tax_rev_gdp), ~ .x > 0)) %>% 
  janitor::adorn_totals("row") %>% 
  knitr::kable(
    col.names = c("Country code", "Country", "Log GDP p.c.", "Log Pop.",
                  "log of Luminosity", "Electric Access", "Tax Revenue % GDP"),
    caption = "Missing observations by country for key covariates")

# Save as rds
write_rds(panel, "data/output/MAIN_panel_data.rds") # Used for all subsequent analyses to avoid re-merging SPI data each time.
# save as csv for AI 
write_csv(panel, "data/output/MAIN_panel_data.csv")
```

## Checking distribution of key variables after log transformation & imputation
```{r}
# checking distribution of key variables after log transformation
key_vars <- panel %>%
  select(spi_overall, log_gdppc, log_pop, electric_access, log_luminosity, rural_pop_pct) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "value")

ggplot(key_vars, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ variable, scales = "free") +
  theme_minimal() +
  labs(title = "Distribution of Key Variables", x = "Value", y = "Frequency")
```


## Quick check of multicollinearity and missingness in control variables
```{r}
# VIF 
vif_data <- panel %>%
  select(prop_sdg_missing, lib_dem, log_gdppc, log_pop, log_luminosity,
         electric_access, rural_pop_pct, spi_overall) %>% na.omit() 

# VIF 1 (with SPI)
vif_model <- lm(prop_sdg_missing ~ lib_dem + log_gdppc + log_pop + electric_access + log_luminosity + rural_pop_pct + spi_overall, data = vif_data)
vif(vif_model)

# VIF 2 (without SPI)
vif2_model <- lm(prop_sdg_missing ~ lib_dem + log_gdppc + log_pop + electric_access + log_luminosity + rural_pop_pct, data = vif_data)
vif(vif2_model)

# matrix of variables and correlations
cor_data <- vif_data %>% select(-prop_sdg_missing)
cor_matrix <- cor(cor_data)
cor_matrix %>% round(2) %>% knitr::kable(
  caption = "Correlation matrix for control variables (excluding SPI)")

```
NOTES ABOUT VARIABLES: 
- Took out R&D expenditure because it has a lot of NAs and is highly correlated with GDP per capita.
- SPI is missing all of 2015, but is kept in the dataset for now because it is a key control variable. Models are run with and without it to check robustness.

## Robustness Check: Models with and without Log of GDP per capita 
```{r}
source("causal_inference_mods/RobChk_GDP_ElecAccess.R") 

# All Three Models (Overall Missingness) in One Table
etable(mA, mB, mC)

# full coefficient table (per goal)
coef_ABC
```


# APPROACH 1.1: Two‑way fixed effects (TWFE) [Baseline Models]
```{r}
# ---------------------- aut_ep -> overall SDG missingness ---------------------

# M1: TWFE WITHOUT SPI OR CONTROLS (just to see raw relationship)
m1_twfe <- feols(
  prop_sdg_missing 
  ~ aut_ep 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m1_twfe)

# M2: TWFE WITH SPI, WITHOUT CONTROLS
m2_twfe <- feols(
  prop_sdg_missing 
  ~ aut_ep 
  + spi_overall 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)

summary(m2_twfe)

# M3: TWFE WITHOUT SPI, WITH CONTROLS
m3_twfe <- feols(
  prop_sdg_missing 
  ~ aut_ep 
  + log_pop 
  + electric_access + I(electric_access^2)
  + log_luminosity
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)
summary(m3_twfe)

# M4: TWFE WITH SPI AND CONTROLS (main model)
m4_twfe <- feols(
  prop_sdg_missing 
  ~ aut_ep 
  + spi_overall 
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  cluster = ~ country_code
)
summary(m4_twfe)
```

# APPROACH 1.2: Two‑way fixed effects (TWFE) [+ Non-Linear Form Terms]
```{r}
# (A) Dynamic TWFE with lagged treatment [BEST MODEL]
mA_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1)
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code)

summary(mA_twfe)

# (B) Dynamic + Income heterogeneity: Ability vs willingness: aut_ep × income_level
mB_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1) * income_level 
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code)

summary(mB_twfe)

# (C) dynamic + SPI heterogeneity (appendix): Ability vs willingness: aut_ep × income_level
mC_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1) * spi_overall 
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code)

summary(mC_twfe)

# (D) dynamic + regime heterogeneity (appendix)
mD_twfe <- feols(
  prop_sdg_missing 
  ~ l(aut_ep, 0:1) * regime_type_4 
  + log_pop 
  + log_luminosity
  + electric_access + I(electric_access^2) 
  + rural_pop_pct 
  | country_code + year,
  data = panel,
  panel.id = ~ country_code + year,
  cluster = ~ country_code)

summary(mD_twfe)

```

A: Does being in an autocratization episode increase missingness, net of capacity and standard controls?
B: ...
C: Does the effect of backsliding on missingness depend on regime type (regime “willingness” to withhold), conditional on spi_overall?
D: ...


